---
title: "Joint Report on SDPD's Gang Enforcement Practices"
subtitle: "DRAFT"
author: "Catalyst California and Pillars of the Community"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
    toc: yes
    toc_depth: 2
    toc_float: yes
---
  
<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)


list.of.packages <- c("usethis", "highcharter", "tidyverse", "sf", "leaflet", "sp", "htmltools", "magrittr", "flextable", "htmlwidgets")

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

#### Load libraries ####
library(usethis)
library(highcharter)
library(tidyverse)
library(sf)
library(leaflet)
library(sp)
library(htmltools) # for html escape
library(magrittr)
library(flextable)
library(lorem)
library(htmlwidgets)

#### Add database connection ####
source("W:\\RDA Team\\R\\credentials_source.R")

con <- connect_to_db("rjs_pillars")

## Source the Chart Template

source("report_embeds/chart_functions.R")

## Reknit all embeds so git will recognize file changes
# source("report_embeds/run_embeds.R")

peridot <- "#CEEA01"
orange <- "#F25922"

```

*Topline: The SDPD criminalizes communities of color, especially Black people, in Southeast San Diego by falsely labeling them as gang members.  This routine practice inflicts devastating racially biased harms, undermines community safety, and wastes tremendous public dollars.*  

<p style="background-color:#CEEA01;font-size:14pt">Community quote or video
</p>

# ACKNOWLEDGEMENTS

`r lorem::ipsum(paragraphs = 1,sentences=NULL, avg_words_per_sentence=10)`


# CONTEXT

Law enforcement agencies throughout California have a long history of unjustly criminalizing people of color—especially  those who are Black and/or Latinx—by documenting them as a member or affiliate of a gang and placing their identities in electronic gang databases. In 1987, for instance, the Los Angeles Sheriff’s Department and the Law Enforcement Communication Network (NALECOM) – a rapid interstate law enforcement communications system^[https://www.ojp.gov/ncjrs/virtual-library/abstracts/national-law-enforcement-communication-network-nalecom-users] – jointly developed a database used to collect, retain, and analyze information about alleged gang members.^[See Tracked & Trapped at 2.] Over the next decade, law enforcement agencies throughout California refined and expanded gang databases.^[Id. at 3.]

In 1997, a company called CSRA International, Inc. (“CSRA”) used funding from the California Department of Justice to develop the most prominent system – “CalGang” (also known as “GangNet”). CalGang allows law enforcement throughout California to collect, store, and share personal information about people suspected – but not convicted – of being involved in criminal activity.^[State Auditor at 2.] People added to CalGang are documented as either a gang member or gang affiliated.^[Id. at 11.] In addition to in-state law enforcement agencies, CalGang is also used by federal law enforcement, other states, and Canada.^[Id. at 9.] Reports indicate that in 2022, at least 106 law enforcement agencies used CalGang, including the San Diego Police Department (SDPD).      

[Insert Table SDPD 2022 CalGang Records Summary (from DOJ Annual Report)]

Research shows that many people have been added to gang databases without prior knowledge and without being arrested for any crime.  Officers often conduct pretextual stops of Black and Latinx people, interrogate them, and document their personal information on a Field Interview (FI) card.  The type of information collected may include, but not be limited to, the person’s name, date of birth, race, gender, nicknames, their family members and friends, their place of residence, the community where they live, photos, tattoos, clothing color, vehicles, and suspected gang affiliation.^[Tracked & Trapped at 4-5. State Auditor at 11.]

<p style="background-color:#CEEA01;font-size:14pt">Community quote or video
</p>

<p style="background-color:#F25922;font-size:14pt">Criminalization callout box: This is “Criminalization” (define via callout box) and it Inflicts Devastating Racially Biased Harms (direct and indirect)</p>

<p style="background-color:#CEEA01;font-size:14pt">Community quote or video
</p>

# FINDINGS

## 1. SDPD targets communities of color, particularly Black and young people in Southeast San Diego, with this harmful practice

In the Southeast community, SDPD’s practice of racial profiling people for gang activity is far too common and affects the daily lives of many in the Southeast community. Both community stories and analysis of police data at various levels of SDPD reveal the racial disparities and harm caused by these practices.

<p style="background-color:#CEEA01;font-size:14pt">Community quote or video
</p>

### A. Data show that stops made by the SDPD are racially biased.

Regardless of whether stops are done under the guise of patrolling for gang activity, officers across SDPD divisions and assignments use racial profiling in their practices. An analysis of all stops voluntarily initiated by officers, referred to as officer-initiated stops, shows that officers disproportionately stop people of color, particularly people officers perceive as Black or Native Hawaiian or Pacific Islander (NHPI). In 2022, for every thousand Black people in San Diego, officers stopped nearly 216 people they perceived as Black, or a total of 16,874 people. In comparison, the average stop rate in San Diego in 2022 was 64 people for every thousand people in the city. Meaning, officers are more than three times as likely to stop a Black person in San Diego than city residents at large. And while Black people comprise only 5.7 percent of the total population in San Diego, they accounted for 19.2 percent of officer-initiated stops.

```{r bubble-chart, child= 'report_embeds/report_stoprates_race_person_bubble.Rmd', echo=FALSE}
```

An assumption often made is officer stops are a means of increasing safety. However, data repeatedly show that police stops are driven by racial bias rather than safety risks. The percentage of stops that result in no action or minor results, such as a warning, is often used as an indicator of racial bias. These stops, known as pretextual stops, show evidence that an officer initiated a stop on a “hunch” that a person was connected to illegal activity but found no serious safety risk once the stop was made. They are consistently found to be used against people of color, exposing them to unnecessary contact with armed law enforcement. 

In San Diego, stops officers choose to initiate against groups of color are more likely to show evidence of pretextual stops. For Black and Latinx people, officer-initiated stops are often more likely to result in no action or a warning. In 2022, nearly 1 in 4 stops (25.3%) of people officers perceived as Black resulted in a warning and over 1 in 10 (12.7%) resulted in no action. Combined these stops affected an estimated 6,416 Black people in San Diego. People officers perceived as Latinx have the second highest rate of stops resulting in no action (11.5%) or a warning (26.8%). These racially biased stops affected an estimated total of 11,563 Latinx people.


```{r bubble-chart, child= 'report_embeds/report_stoprates_result_race_person_stacked.Rmd', echo=FALSE}
```

### B. Available data show that SDPD’s gang enforcement assignment profiles Black and Latinx people for gang activity.   

Stories from community members indicate the practices of SDPD’s Gang Suppression Unit are even more troubling. Gang enforcement officers target and harass Black and Latinx people and the Southeast community, unfairly labeling them as gang members. 

<p style="background-color:#CEEA01;font-size:14pt">Community quote or video
</p>

Limited RIPA data on officers assigned to the gang enforcement assignment at the time of the stop demonstrate how these officers specifically target Latinx and Black people. Out of people stopped by officers assigned to gang enforcement, 40.1% are people officers perceived as Latinx and 32.4% are people perceived as Black. Compared to their share of the San Diego population, Black and Latinx people are 5.7 and 1.3 times overrepresented in people stopped by officers assigned to gang enforcement.  

Black people of every age group in San Diego are disproportionately impacted by these stops. For instance, while Black people ages 45-54 comprise just 5.2% of the San Diego population in that age group, they made up 41.5% of people stopped by officers assigned to gang enforcement. This translates into Black people ages 45-54 being nearly 8 times overrepresented in people stopped by officers assigned to gang enforcement. Latinx young to middle-aged adults are similarly overrepresented. Latinx people between the ages of 18-44 are over 1.3 times overrepresented in people stopped by officers assigned to gang enforcement compared to their share of the San Diego people. While RIPA data only includes a limited number of stops made by the gang enforcement assignment, these trends indicate the alarming rate of racial profiling used by the Gang Suppression Unit. 

```{r bubble-chart, child= 'report_embeds/report_stoprates_race_gang_person_bubble.Rmd'}
```


<div style="background-color:#F25922">
### SDPD data are underinclusive and mask the extent of racially biased, gang profiling in San Diego.

Research suggests that SDPD data on gang documentation are underinclusive and, in turn, mask scope of racial biases because of a lack of clarity (and inconsistencies) between the officer assignment categories used under RIPA regulations, SDPD’s list of units, and the functional tasks SDPD’s officers are responsible for irrespective of their assignments. 

More specifically, RIPA regulations require that the “type of assignment of officer” be reported for each stop.  It includes eleven options, including “gang enforcement.”^[11 Cal. Code Regs. § 999.226(a)(20)]  Research indicates that SDPD has approximately 39 departmental units to which officers are assigned.  This includes a “gang unit” for personnel that, among other things, identify and investigate gang members and gang activity, maintain SDPD gang information files and liaise with agencies across the nation, and assist patrol with gang information.^[https://www.sandiego.gov/sites/default/files/legacy/police/pdf/unitlist.pdf]  This latter function, that is, assisting patrol with gang information, indicates that the gang unit collaborates with other parts of SDPD that conduct patrol activities (such as traffic-related units) on matters related to alleged gang activity that may not qualify as “gang enforcement” under RIPA’s regulation on officer assignment.

To account for this, two public records requests were submitted to SDPD to determine whether there were additional stops related to gang enforcement that may not have been reported under RIPA’s gang enforcement assignment due to the lack of clarity between RIPA and SDPD definitions and personnel functions.  Both requests were denied.^[insert brief description of SPDP’s responses]  SDPD’s failure to be transparent with RIPA data on stops masks the extent of racial biases in SDPD’s gang documentation activities and, in turn, attendant harms that are disproportionately experienced by Black and Latinx community members in Southeast San Diego.   

Because SDPD refused to provide data specific to the Gang Suppression Unit, we use these stops that result in a field interview card as a proxy in detailing gang suppression activity from SDPD. We combine this data with stories from the community to show the extent of this harmful practice.
</div>

### C. Beyond the gang enforcement assignment, an analysis of stops that result in field interview cards also show that SDPD targets Black and Latinx people, unfairly labeling them as gang members. 

Community members’ experiences detail how SDPD’s gang profiling activity subjects them to invasive and traumatic encounters that are underinclusive in what SDPD reports. These encounters often include field interview cards that are used to enter community members into California’s Cal Gang database. 

<p style="background-color:#F25922;font-size:14pt">FI Card explanation/definition</p>


<p style="background-color:#CEEA01;font-size:14pt">Community quote or video
</p>

Data shows that SDPD stops people they perceive as Black for field interview cards at the highest rate compared to all other racial groups. Despite Black people making up only 5.6% of San Diego’s total population, the rate at which SDPD stops people they perceive as Black for field interview cards is 5 times more than the rate at which they stop people perceived as White. Meanwhile, White people make up 42% of San Diego city’s total population.  

Communities most impacted by police gang enforcement activity report that SDPD targets youth of color for gang activity. The data reflects this trend and shows alarming disparities between all other age groups. Young adults ages 18-24 and minors 17 years old and younger are more likely to be stopped and have a field interview card completed compared to all other age groups. But, regardless of age, SDPD stops and completes field interview cards for individuals they perceive as Black or Latinx at higher rates compared to other racial groups. 

```{r bubble-chart, child= 'report_embeds/report_stoprates_ficard_race_person_bubble.Rmd'}
```

SDPD’s racial targeting for stops that result in a field interview card primarily takes place in South and Southeast San Diego. Across all SDPD police divisions, SDPD makes the most amount of stops that result in a field interview card in the Central, Southeastern, and Mid-City divisions. In 2022, SDPD made 1,100 total stops with a field interview card in the Southeastern division compared to only 171 total stops resulting in a field interview card in the Northwestern division, an affluent community in San Diego with a higher percentage of White people. Analysis by race shows that the people SDPD disproportionately stops for field interview cards are people of color, especially people perceived as Black and Latinx. In the Southeastern division, 1 in 5 people (19.4%) SDPD stops and perceives as Black are only stopped for field interview cards compared to less than 1 in 10 White people they stopped.

```{r dot-density, child= 'report_embeds/report_stoprates_div_ficard_race_person_density_map.Rmd'}
```

Despite the high rate of people of color stopped with field interview cards completed, the data shows that SDPD makes these stops based on pretext and racial profiling, rather than risk for crime. Nearly 1 in 4 stops that result in a field interview card are originally made for traffic violations. Analyzing those traffic stops by race show that SDPD disproportionately use the guise of traffic violations to stop and complete field interview cards for individuals they perceive as people of color compared to people perceived as White. Individuals perceived as people of color are stopped for a traffic violation and have field interview card completed at more than twice the rate of people who are perceived as White by SDPD. While the officer reportedly made these initial stops in response to a traffic safety risk, they concluded in no result indicative of an actual traffic crime being completed, such as a citation. Instead, the officer only completed a field interview card, indicating pretext and the use of racial profiling to harass people of color for gang activity.

```{r stacked-chart, child= 'report_embeds/report_stoprates_ficard_reason_race_bubble.Rmd'}
```

## 2. Such racially biased stops harm the Southeast community and undermine community safety.  

SDPD’s gang profiling activity subjects community members to unnecessary contact with armed officers and undermines the public and individual safety of the community. Stories from community members detail the harm caused by SDPD’s gang enforcement. For instance, community members are repeatedly exposed to physical searches by officers that traumatize them and lead to feeling a lack of safety from the institution chartered to protect them.

<p style="background-color:#CEEA01;font-size:14pt">Community quote or video
</p>

### A. When searches are conducted as a part of gang profiling, they are largely ineffective and biased.  

During an officer-initiated stop, an officer may decide to search a person with the assumption that the person is carrying contraband or evidence of a crime. Data on police stops repeatedly show that officer searches are highly ineffective in rooting out crime and are racially biased. Out of all the people officers chose to search in 2022, officers failed to find evidence or contraband on 70 percent of them. In other words, over 18,000 people were subjected to a police search by SDPD in 2022 that was proven unnecessary.

```{r donut-chart, child= 'report_embeds/report_hitrates_race_person_donut.Rmd'} 
```

This is also the case for searches done during field interview card stops. When officers conduct searches during these stops, they show greater evidence of racial profiling. Officers failed to find contraband or evidence approximately 72 percent of the time when conducting a search during a field interview card. In 84 percent of searches done during stops where a field interview card was completed of individuals perceived as NHPI, SDPD officers found no contraband or evidence. This rate drops to 65 percent during field interview card stops of people perceived as White, showing that searches for contraband are more ineffective when applied to people of color. These unnecessary searches and contact with officers affected the highest numbers of Latinx and Black people. In 2022, officers subjected 1,268 people perceived as Latinx and 995 people perceived as Black to these searches.

```{r stacked-chart, child= 'report_embeds/report_hitrates_race_ficard_person_bubblepop.Rmd'}
```

### B. Southeast community members have had their wellbeing and safety harmed by gang enforcement stops.

While these SDPD practices occur throughout San Diego, data and stories from the Southeast community show the prevalence of these searches in Southeast San Diego. Community members describe how these searches are overwhelmingly prevalent in their neighborhoods and expose them to contact with police that unduly creates trauma and harm for the community.

<p style="background-color:#CEEA01;font-size:14pt">Community quote or video
</p>

Analysis of searches done during field interview card stops further demonstrates the extent of this harmful practice and how it targets people of color in San Diego, and particularly, the Southeast community. In all of San Diego, there were 2,936 people who were searched, field interviewed, and found to have no contraband. Of these people, around 66 percent were in the Southeastern, Central, and Mid-City divisions of San Diego, areas that have a higher concentration of people of color in San Diego compared to the Northern regions of San Diego.  

The Southeast Division has the highest number of people searched during field interview card stops and the highest number of people found with no contraband or evidence. In 77 percent of searches done during stops where a field interview card was completed in the Southeastern Division for all individuals, SDPD officers found no contraband or evidence. Black and Latinx people in the Southeast are disproportionately targeted by these searches. Individuals perceived as Black and Latinx accounted for over 89 percent of these unsuccessful searches while only making up 65 percent of the population in this division. These disparities exist in other divisions as well. While Black people represent only seven percent of the population in the Central division, they comprised 35 percent of people searched with no contraband or evidence. And while they represent the second highest searched group in the Central Division, only second to Latinx individuals, officers failed to find contraband or evidence on nearly 81 percent of people they perceived as Black who they searched and field interviewed.

```{r dot-map, child= 'report_embeds/report_hitrates_div_race_ficard_person_density_map.Rmd'}
```

## 3. These racially biased activities waste public resources.

Like most police departments in California, a significant portion of public dollars goes to SDPD. The narrative pushed by police departments is that a huge amount of funding is needed to keep communities safe. However, data shows that police spend their time on activities that are not preventing crime or keeping the public safe. SDPD is no exception to this.

<p style="background-color:#CEEA01;font-size:14pt">Community quote or video
</p>

### A. Based on available data, SDPD spends a disproportionate amount of time on profiling people of color for gang activity.

When analyzing how much time SDPD spends on stops resulting in a field interview card, nearly six percent of all hours spent on officer-initiated stops are spent on stops that result in a field interview card.  

Within stops that result in a field interview card, SDPD disproportionately spends their time stopping people perceived as Latinx and Black. More than half of the hours spent on stops that result in a field interview card are spent stopping people perceived as Latinx or Black though Latinx and Black people make up only 35.7 percent of the population in San Diego.

```{r item-chart, child= 'report_embeds/report_timespent_ficard_race_stop_item.Rmd'}
```

Analysis of hours spent by SDPD officers assigned to gang enforcement also show that most of their time is not being spent on activities that prevent crime. Offices assigned to gang enforcement spent nearly 4 in 10 hours on stops resulting only in a field interview card, a warning, or no action. Stops only resulting in a citation for an infraction took up an additional 8 out of 100 patrol hours. Comparatively, arrests with warrants only comprised 9 in 100 patrol hours.

```{r item-chart, child= 'report_embeds/report_timespent_ficard_gang_stop_item.Rmd'}
```

### B. This time translates to wasted public dollars.

`r lorem::ipsum(paragraphs = 1,sentences=NULL, avg_words_per_sentence=10)`


# RECOMMENDATIONS

1. Alternative investments that actually increase safety
2. Enact the Protect Act
3. Others...


# APPENDIX

```{r stacked-chart, child= 'report_embeds/report_stopsrates_age_race_gang_person_stacked.Rmd'}
```


```{r stacked-chart, child= 'report_embeds/report_stoprates_age_race_ficard_person_stacked.Rmd'}
```


# METHODOLOGY

[link to methodology on GitHub]