---
title: "Joint Report with Pillars of the Community on SDPD Gang Enforcement"
subtitle: "DRAFT"
author: "Catalyst California and Pillars of the Community"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
    toc: yes
    toc_depth: 2
    toc_float: yes
---
  
<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)


list.of.packages <- c("usethis", "highcharter", "tidyverse", "sf", "leaflet", "sp", "htmltools", "magrittr", "flextable", "lorem",  "htmlwidgets", "vembedr", "knitr")

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

#### Load libraries ####
library(usethis)
library(highcharter)
library(tidyverse)
library(sf)
library(leaflet)
library(sp)
library(htmltools) # for html escape
library(magrittr)
library(flextable)
library(lorem)
library(htmlwidgets)
library(vembedr)
library(knitr)

#### Add database connection ####
source("W:\\RDA Team\\R\\credentials_source.R")

con <- connect_to_db("rjs_pillars")

## Source the Chart Template

source("report_embeds/chart_functions.R")

## Reknit all embeds so git will recognize file changes
# source("report_embeds/run_embeds.R")

peridot <- "#CEEA01"
orange <- "#F25922"

```

**The San Diego Police Department criminalizes communities of color, especially Black and Latinx people, in Southeast San Diego through gang profiling. This practice inflicts devastating racially biased harms, undermines community safety, and wastes tremendous public dollars.** 

```{r}
embed_url("https://youtu.be/cWlR-a1bm5w")

```

# SUMMARY

This report combines data analysis of San Diego Police Department’s (“SDPD”) patrol activities, video interviews with community members in Southeast San Diego, and public policy research to show how SDPD criminalizes community members in Southeast San Diego through gang profiling. Gang profiling is a practice through which SDPD–and law enforcement agencies throughout California–document people of color as gang members and place their identities in gang databases. This report unpacks how gang profiling in Southeast San Diego inflicts devastating racially biased harms, undermines community safety, and wastes tremendous public dollars. To address these issues, a comprehensive set of recommendations for the City of San Diego (the “City”) are provided. 


# ACKNOWLEDGEMENTS

This report was jointly produced through a partnership between Catalyst California and Pillars of the Community–an organization that advocates and organizes for people harmed by the criminal legal system in Southeast San Diego. Thank you to Southeast San Diego community members for courageously sharing your stories and perspectives on how SDPD patrol activities undermine community safety. This project was completed with support from the Microsoft Justice Reform Initiative. 


# BACKGROUND

The SDPD has a long history of criminalizing people of color in Southeast San Diego through a devastating practice known as “gang profiling.” Used by California law enforcement agencies for decades, gang profiling occurs through officers documenting community members–especially those who are Black and/or Latinx–as a member or affiliate of a gang and placing their identities in electronic gang databases. For example, in 1987, the Los Angeles Sheriff’s Department (“LASD”) and the Law Enforcement Communication Network (“NALECOM”)–a rapid interstate law enforcement communications system^[https://www.ojp.gov/ncjrs/virtual-library/abstracts/national-law-enforcement-communication-network-nalecom-users]–jointly developed and began using a database to collect, retain, and analyze information about alleged gang members.^[See Tracked & Trapped at 2.] Over the next decade, gang databases were refined and expanded.^[Id. at 3.]

In 1997, the most prominent gang database–CalGang (also known as “GangNet”)–was developed through funding from the California Department of Justice. CalGang allows law enforcement throughout California to collect, store, and share personal information about people suspected, but not convicted, of being involved in criminal activity.^[State Auditor at 2.] People added to CalGang by law enforcement are documented as either a gang member or gang affiliated.^[Id. at 11.] In addition to in-state law enforcement agencies, CalGang is used by federal law enforcement, other states, and Canada.^[Id. at 9.] Reports indicate that in 2022, at least 106 law enforcement agencies used CalGang, including SDPD. Amongst those agencies, SDPD had the third highest number of records in CalGang (only LASD and the San Bernardino County Sheriff’s Department had more). Latinx and Black people respectively accounted for 46% and 43% of people added to CalGang by SDPD.    

```{r calgang-table, child= 'report_embeds/report_calgangstats_table.Rmd', echo=FALSE}
```

The process through which people of color are gang profiled is notoriously problematic. Many are added to gang databases without prior knowledge and without being arrested for a crime. For example, officers often conduct pretextual stops of Black and Latinx people by detaining them for relatively minor issues—such as a broken taillight or jaywalking—and interrogating them in aim of discovering evidence of a more serious crime. Through the interrogation, which typically functions as a form of intimidation and harassment, officers document personal information about the community member on a Field Interview card. The information collected may include, but not be limited to, the person’s name, date of birth, race, gender, nicknames, names of family members and friends, their place of residence, the community where they live, photos, images and descriptions of their tattoos, clothing colors at the time of a stop, vehicles registered to or driven by the person stopped, and suspected gang affiliation identified by the officer.^[Tracked & Trapped at 4-5. State Auditor at 11; See also https://gould.usc.edu/students/journals/rlsj/issues/assets/docs/issue_18/Yoshino_(MACRO2).pdf] Being documented as gang members imposes devastating, long-lasting harms on Southeast San Diego community members.

```{r}
embed_url("https://www.youtube.com/watch?v=FTqt90YJ_V4")

```

<div style="background-color:#F25922">
## The Devastating Impact of Criminalizing People through Gang Profiling
Criminalization occurs when people are turned into criminals by making their alleged activities illegal–including actions that are ordinary in a community, not harmful to others, or byproducts of systemic racism.^[https://www.oed.com/search/dictionary/?scope=Entries&q=criminalization+ ; https://www.merriam-webster.com/dictionary/criminalize] SDPD criminalizes Southeast San Diego community members by documenting them as gang members. This inflicts devastating harms–including physical, mental, and emotional trauma. 

In addition, being documented as a gang member has devastating ripple effects, especially for people of color. For example, a person charged with an alleged crime can be subject to longer periods of incarceration if a sentence enhancement for gang membership (i.e., a “gang enhancement”) is applied. Specifically, the application of a gang enhancement can add an additional two years to life imprisonment to a person’s sentence. The added time runs consecutive (rather than concurrent) to the punishment for the underlying felony.^[Cal. P.C. s. 186.22(b); Yoshino at 119.] These factors have devastating impacts on people facing relatively low charges because the threat of far more lengthy imprisonment impacts plea bargaining, charging decisions, juror perspectives, and other factors in criminal cases.^[Yoshino at 138.] For immigrant communities, this often includes potential deportation.^[https://link.springer.com/article/10.1007/s10612-022-09614-3 ] As context for the racially inequitable impact of gang profiling, statewide data show that “Black and Latinx people comprise 92% of the people sentenced under California’s gang enhancement statute”^[http://www.clrc.ca.gov/CRPC/Pub/Reports/CRPC_AR2020.pdf at 45.] while collectively accounting for approximately 45% of the state population.^[https://www.ppic.org/publication/californias-population/#:~:text=No%20race%20or%20ethnic%20group,the%202022%20American%20Community%20Survey.]   
</div>


# BRIEF METHODOLOGY

This report evaluates SDPD gang profiling by analyzing 2022 data collected and reported by SDPD pursuant to the Racial and Identity Profiling Act (“RIPA”) of 2015. For each stop made by law enforcement, RIPA data includes information that can be analyzed for evidence of profiling—including characteristics about the person stopped (e.g., race, gender, and age), the stop location, the length of the stop, the reason for the stop, and the result of the stop. This report uses RIPA data to evaluate SDPD’s gang profiling activities. Key RIPA indicators include the officer assignment field (under “gang enforcement”), and the stop result field (under “field interview card”). 


# KEY FINDINGS

In sum, stop data, video stories from community members, and policy research show that SDPD is committed to a practice of criminalizing communities of color–especially Black and Latinx people–in Southeast San Diego through gang profiling. This causes severe physical, mental and emotional harm, and operates as a form of dehumanization. Below are key findings.

1. SDPD patrol activities target Black and Latinx people (and in some instances other people of color) in Southeast San Diego and subject them to gang profiling.     

2. SDPD’s gang profiling undermines community safety by inflicting tremendous harm on Southeast San Diego community members while being an ineffective crime prevention method.     

3. SDPD’s racially biased gang profiling in Southeast San Diego wastes tremendous time and public dollars. Those resources could be better invested in non-law enforcement safety programs that are rooted in care and equity.     

The data and stories below unpack these findings in greater detail.


## 1. SDPD targets communities of color, particularly Black and Latinx people in Southeast San Diego, with gang profiling.

Data below show that stops made by SDPD are racially biased, and officers assigned to gang enforcement disproportionately profile Black and Latinx people. Beyond officers assigned to gang enforcement, an analysis of field interview cards reveals racially biased gang profiling. And, community member stories translate the data into lived experiences that reveal unsettling harm and trauma caused by SDPD.

```{r}
embed_url("https://www.youtube.com/watch?v=arZekGIgIrw")

```

### A. Stops made by SDPD are racially biased.

An analysis of all officer-initiated stops^[As opposed to stops based on calls for service from community members.] shows that officers disproportionately stop Black, Native Hawaiian or Pacific Islander (NHPI), and Latinx people. More specifically, in 2022, for every 1000 Black people in San Diego, officers stopped nearly 216 people they perceived as Black. To compare, the citywide average stop rate was 64 people stopped for every 1000 people. This means that **SDPD was more than three times more likely to stop a Black person than residents at large.** In terms of percentages, Black people comprised 5.7 percent of the total population, but accounted for 19.2 percent of officer-initiated stops.

```{r bubble-chart, child= 'report_embeds/report_stoprates_race_person_bubble.Rmd', echo=FALSE}
```

In addition, data  show that officer-initiated stops are often driven by racial bias, not safety risks. Specifically, stops that result in “no action” or a relatively minor result (such as a warning) indicate relatively low safety risk. When people of color are disproportionately subjected to stops that pose little-to-no safety risk, it indicates racial bias. These stops, known as “pretextual,” are often based on a racially biased and incorrect “hunch” that a person of color is connected to illegal activity. All too often, pretextual stops expose people of color to unnecessary contact with the criminal legal system. They are a primary method of racial profiling. 

With this context, data show that SDPD disproportionately subjects Black and Latinx people to officer-initiated stops that result in no action or a warning. In 2022, nearly one in four stops (25.3%) of people officers perceived as Black resulted in a warning, and over one in ten (12.7%) resulted in no action. Combined these stops affected an estimated 6,416 Black people. For people that officers perceived as Latinx, 11.5% of officer-initiated stops resulted in no action and 26.8% resulted in a warning (26.8%). Those stops directly affected approximately 11,563 Latinx people that year.

```{r bubble-chart, child= 'report_embeds/report_stoprates_result_race_person_stacked.Rmd', echo=FALSE}
```

### B. SDPD gang enforcement officers target Black and Latinx people for profiling.   

Stories from community members reveal that SDPD targets and harasses Black and Latinx people, especially in Southeast San Diego, by labeling them as gang members. In addition, officers often target youth of color who pose no threat to safety.

```{r}
embed_url("https://www.youtube.com/watch?v=Dd-e07UT6-Y")

```

Data show that officers assigned to gang enforcement under RIPA data target Latinx and Black people. Specifically, amongst all people stopped by officers assigned to gang enforcement, 40.1% were perceived as Latinx and 32.4% were perceived as Black. Compared to their proportions of the San Diego population, **Black and Latinx people were respectively 5.7 and 1.3 times overrepresented among people stopped by officers assigned to gang enforcement.** 

In addition, Black people in every age group were disproportionately impacted. For instance, while Black people comprised 5.2% of the 45 to 54 years old population, they accounted for 41.5% of people in that age group who were stopped by officers assigned to gang enforcement. Stated differently, Black people ages 45 to 54 were nearly eight times overrepresented among people stopped by officers assigned to gang enforcement. Latinx people between the ages of 18 and 44 were more than 1.3 times overrepresented among people stopped by officers assigned to gang enforcement compared to their proportion of the population. Notably however, as explained below, such figures likely undercount the number of stops conducted by gang enforcement.

```{r bubble-chart, child= 'report_embeds/report_stoprates_race_gang_person_bubble.Rmd'}
```


<div style="background-color:#F25922">
### SDPD data are underinclusive and mask the extent of racially biased gang profiling in San Diego.

Research suggests that publicly available SDPD data on gang documentation are underinclusive and, in turn, mask the scope of racial biases in SDPD gang enforcement. This is in part due to a lack of clarity (and inconsistencies) between the officer assignment categories used under RIPA regulations, SDPD’s list of units, and the functional tasks SDPD’s officers are responsible for irrespective of their assignments. 

Specifically, RIPA regulations require that the “type of assignment of officer” be reported for each stop.^[11 Cal. Code Regs. § 999.226(a)(20)] There are 11 assignment types, including “gang enforcement.” Catalyst California’s research indicates that SDPD has approximately 39 departmental units to which officers are assigned. This includes a “gang unit” for personnel that, among other tasks, identify and investigate gang members and gang activity, maintain SDPD gang information files and liaise with agencies across the nation, and assist patrol with gang information.^[https://www.sandiego.gov/sites/default/files/legacy/police/pdf/unitlist.pdf] This latter function—assisting patrol with gang information—indicates that the gang unit collaborates with other parts of SDPD that conduct patrol activities (such as traffic-related units) on matters related to alleged gang activity that may not qualify as “gang enforcement” under RIPA’s regulation on officer assignment.

To account for this, Catalyst California submitted a public records request to SDPD to determine whether there were additional stops related to gang enforcement that may not have been reported under RIPA’s gang enforcement assignment due to the lack of clarity between RIPA data and SDPD functions. The request and an appeal were denied.^[insert brief description of SPDP’s responses] SDPD’s lack of transparency masks the extent of racial biases in SDPD’s gang profiling and, in turn, attendant harms that are disproportionately experienced by Black and Latinx community members in Southeast San Diego. 

Because SDPD refused to provide data specific to its gang unit, available data on stops that result in a field interview card were analyzed as a proxy for identifying gang profiling activities. This data, combined with stories from impacted community members, further reveal the extent of SDPD’s extremely harmful gang profiling.

</div>


### C. Beyond officers assigned to gang enforcement, an analysis of stops that result in field interview cards shows that SDPD targets Black people for gang profiling, especially in Southeast San Diego. 

Community member experiences vividly reveal how SDPD’s gang profiling subjects people to trauma and harassment that goes far beyond what data show. These experiences include intrusive interrogations used to complete field interview cards (a key indicator of gang profiling) so that SDPD can document people in Southeast San Diego as gang members and add them to CalGang. 

```{r}
embed_url("https://www.youtube.com/watch?v=oOmzLOEOFGM")

```

Compared to all racial groups, data show that SDPD subjects people they perceive as Black to field interviews at the highest rate. Black people were subjected to stops that resulted in a field interview card at a rate five times greater than White people.

```{r bubble-chart, child= 'report_embeds/report_stoprates_ficard_race_person_bubble.Rmd'}
```

Analyzing data by SDPD’s patrol divisions shows that the department’s practice of disproportionately subjecting people of color to gang profiling primarily takes place in South and Southeast San Diego–areas where more people of color live. Specifically, SDPD stops the highest percentage of people for field interviews in the Central and Southeastern Divisions. **In 2022, SDPD subjected 1,100 people to stops resulting in field interviews in the Southeastern Division (e.g., Lincoln Park, Skyline, etc.) compared to only 171 people in the Northwestern Division.^[In the Southeastern Division this equated to 13.8% of people stopped compared to just 5% of people stopped in the Northwestern division.] More White and affluent people reside in the Northwestern Division**, which encompasses areas like Carmel Valley, Del Mar Heights, and parts of Torrey Pines. **Black people comprise less than 1% of the population in the Northwestern division, but account for 14.9% of the population in Southeast San Diego.**

Analysis by race shows that SDPD disproportionately subjects people of color to field interviews (which often amount to interrogation and harassment) during stops. For example, **in the Southeastern Division, one in five people (19.4%) that officers stopped and perceived as Black had field interview cards completed, compared to less than one in 10 (9.6%) people perceived as White.**^[Similarly, in the Central Division, 34% of people officers stopped and perceived as Multiracial had field interview cards completed compared to 21.4% of White people.]

```{r dot-density, child= 'report_embeds/report_stoprates_div_ficard_race_person_density_map.Rmd'}
```

In addition, the data indicate that SDPD’s practice of disproportionately stopping people of color and subjecting them to field interviews is rooted in racial profiling. Specifically, nearly one in four stops that resulted in a field interview card originated based on alleged traffic violations. Analyzing those stops by race shows that SDPD used traffic violations to stop people they perceived as people of color for field interview cards at more than twice the rate of people perceived as White. To add, while these stops were allegedly made in response to traffic safety risks, they concluded in results that indicate otherwise because field interview cards, rather than citations or arrests, were issued. This indicates the use of pretextual stops for racially biased gang profiling. 

```{r stacked-chart, child= 'report_embeds/report_stoprates_ficard_reason_race_bubble.Rmd'}
```

## 2. SDPD’s racially biased gang profiling undermines safety and harms the Southeast San Diego community.  

SDPD’s gang profiling unnecessarily subjects people of color to harmful criminal legal system contact and undermines community safety. Specifically, stories from Southeast San Diego community members show that they are often harassed, interrogated, searched, and subjected to other degrading and dehumanizing treatment by SDPD. This results in community members feeling unsafe from SDPD. 

```{r}
embed_url("https://www.youtube.com/watch?v=e1KxqPIKjeo")

```

In addition to being extremely harmful, data show that SDPD’s gang profiling is extremely ineffective. This is shown by SDPD’s discovery (or “hit) rate–which measures how often evidence of a crime is found when a person is searched by an officer.

### A. SDPD searches conducted during gang profiling are highly ineffective and racially biased. 

During routine stops, officers often search people who they think may possess evidence of a crime. Stop data show that amongst all people searched by SDPD in 2022, officers failed to find evidence of a crime over 70% of the time. Outside the context of law enforcement (e.g., education, public health, or housing), a 70% failure rate in government services that purportedly aim to improve communities would, at minimum, place the practice at issue in question. For SDPD, allowing this practice amounted to over 18,000 people being subjected to unnecessary, and often intrusive and traumatizing, searches in one year.

```{r donut-chart, child= 'report_embeds/report_hitrates_race_person_donut.Rmd'} 
```

To add, searches conducted during stops that resulted in the completion of a field interview card–a proxy for gang profiling–were even less effective. Officers failed to find evidence of a crime in approximately 72.1% of those stops. To add, **people of color were less likely to possess evidence of a crime than White people. Specifically, the discovery rate for searches of White people was 34.9%, while the discovery rates for searches of NHPI, Black, Asian, Latinx, SWANA/SA, and AIAN people ranged between 15.8% and 30.4%.** NHPI people had the lowest discovery rates. However, these racially biased and unnecessary searches affected Latinx and Black people the most. Officers subjected 1,705 people perceived as Latinx and 1,366 people perceived as Black to these searches in 2022.

```{r stacked-chart, child= 'report_embeds/report_hitrates_race_ficard_person_bubblepop.Rmd'}
```

### B. Southeast San Diego community members’ wellbeing and safety are harmed by searches that occur during gang profiling.

Combined, stories from Southeast community members and data reveal the severity and prevalence of unnecessary SDPD searches in Southeast San Diego.  

```{r}
embed_url("https://www.youtube.com/watch?v=KzE0t3U4Sk0")

```

Division-level analysis of searches conducted during stops that result in field interview cards shows that people of color in Southeast San Diego are being targeted. Specifically, citywide, 2,936 people were searched, field interviewed, and found to have no contraband in 2022. Approximately 66.3% of those people were in SDPD’s Southeastern, Central, and Mid-City Divisions–areas that have higher concentrations of people of color compared to northern regions. 

Among stops resulting in field interview cards, people in the Southeastern Division were subjected to the most searches and had the highest number of community members found with no contraband. **SDPD officers found no contraband in 77.5% of searches done during stops where a field interview card was completed in the Southeastern Division. Combined, people perceived as Black and Latinx accounted for 89.8% of those unsuccessful searches, while only making up 65.8% of the population in the Division.** Similar disparities were also present in other divisions.^[For example, while Black people represented only 7.4% of the population in SDPD’s Central Division, they comprised 35.1% of people subjected to a search where no evidence of a crime was found. And while they represented the second highest searched group in SDPD’s Central Division, only second to Latinx people, officers failed to find evidence on nearly 81.1% of people they perceived as Black and searched and field interviewed.] 

```{r dot-map, child= 'report_embeds/report_hitrates_div_race_ficard_person_density_map.Rmd'}
```

## 3. SDPD’s racially biased gang profiling in Southeast San Diego wastes tremendous time and public dollars.

In addition to harming Southeast community members and undermining community safety, SDPD annually wastes time and millions of public dollars on gang profiling. Those resources should be invested in non-law enforcement equity-centered programs that advance true safety.

```{r}
embed_url("https://www.youtube.com/watch?v=sD5libtgXfI")

```

### A. SDPD wastes a significant amount of time profiling people of color for gang activity in Southeast San Diego.

More than half (56.1%) of SPDP hours spent on stops that result in field interviews are spent stopping people perceived as Latinx or Black, even though Latinx and Black people only make up 35.7% of the population. In addition, across all SDPD Divisions, the Department spends more time subjecting Black people to field interviews compared to their share of the population. This is problematic because, as explained above, field interviews often entail intrusive questioning that operates as a form of severe intimidation and harassment. **In SDPD’s Southeastern Division, officers spent three times more hours on field interview cards of Black people compared to their share of the population.**

```{r item-chart, child= 'report_embeds/report_timespent_div_ficard_race_person_group.Rmd'}
```

In addition, an analysis of hours spent by SDPD officers assigned to gang enforcement shows that nearly 40 out of every 100 (39.7%) hours was spent on stops with results indicative of little-to-no safety risk (i.e., resulting in a field interview card, a warning, or no action). To add, stops only resulting in a citation for an infraction (as opposed to a misdemeanor or felony) accounted for 8 out of 100 (7.7%) patrol hours. Collectively, this amounted to officers assigned to gang enforcement wasting nearly half (47.4%) of their time on minor conduct.

```{r item-chart, child= 'report_embeds/report_timespent_ficard_gang_stop_item.Rmd'}
```

### B. SDPD annually wastes over $43 million on gang profiling in Southeast San Diego.

In the adopted budget for fiscal year 2022, the city’s general fund expenditures on SDPD were the largest for any department, totaling $593.3 million. That amount was far more than the combined expenditures on other agencies that help build long-term community safety–$358.4 million. For example, the budget for Park & Recreation was $133.9 million, Transportation was $82.5 million, Libraries was $60.7 million, Environmental Services was $58.8 million, and Economic Development was $22.5 million.^[(See page 97 of https://www.sandiego.gov/sites/default/files/fy22ab_v1generalfundexpenditures.pdf)] 


```{r bar-chart, child= 'report_embeds/report_city_budget_bar.Rmd'}
```


While the City’s budget is not publicly disaggregated by SDPD divisions for fiscal year 2022, in 2021 the City’s Independent Budget Analyst produced an in-depth analysis which showed that the SDPD’s total fiscal year 2021 budget was $568.2 million. Amongst that, SDPD’s patrol operations—the part of the SDPD typically responsible for conducting stops—accounted for $238.6 million in general fund expenditures.^[(Pg. 6 of 2021 In-depth analysis - https://www.sandiego.gov/sites/default/files/iba-report-20-21.pdf)] Within that, the Southeastern Division of SDPD patrol operations–which conducts the vast majority of stops in its area–cost $26.2 million.^[(Id. at A-7).] In addition to the Southeastern Division, SDPD has  three special units for gang profiling, including the Street Gang Unit, Gang Intervention Unit, and Special Operations Unit.^[https://www.sandiego.gov/sites/default/files/iba-report-20-21.pdf] The units cost  $16.8 million.^[(Id. at A-64-66).]

[Insert table with above figures like we had for the statewide report]

In light of the above figures, **the estimated cost of operating the parts of SDPD primarily responsible for gang profiling in Southeast San Diego–that is, the Southeastern Division of Patrol Operations, and the Street Gang, Gang Intervention, and Special Operations Units—amounted to approximately $43.1 million in fiscal year 2021.** This does not include the cost of operating SDPD’s most expensive division—the Central Division of Patrol Operations ($32.4 million)—which has a patrol area that includes neighborhoods often considered part of Southeast San Diego (e.g., Barrio Logan and Logan Heights) that have relatively higher Latinx populations. It also excludes millions spent on support operations, human resources, and other administrative functions.

Thus, a conservative estimate shows that the City’s spending on gang profiling in Southeast San Diego in 2021 ($43.1 million) far exceeded its investment in the Economic Development Department ($23.3 million), which implements citywide business and community development programs.^[https://www.sandiego.gov/sites/default/files/fy21ab_v2econdev.pdf. ] More apt comparisons are the City’s investments in the Storm Water Department ($47.5 million)—which is responsible for preventing flooding like that which displaced over 1,200 San Diegans from their homes in January 2024^[https://www.nbcsandiego.com/news/local/more-than-1200-plus-san-diegans-still-homeless-after-the-great-flood-of-2024/3429816/#:~:text=1%2C200%2Dplus%20San%20Diegans%20lost,22%20%E2%80%93%20NBC%207%20San%20Diego] ^[https://www.sandiego.gov/sites/default/files/fy21ab_v2stormwater.pdf]–, and homelessness prevention ($49.1 million).^[https://www.sandiego.gov/sites/default/files/fy21ab_v1generalfundexpenditures.pdf] The City should stop wasting public dollars on SDPD’s gang profiling because it inflicts devastating racially biased harms and undermines community safety. To build true community safety, those dollars should be reinvested in equity- and care-centered programs discussed in the recommendations below.

```{r}
embed_url("https://www.youtube.com/watch?v=W_j4T1uOQhY")

```


# RECOMMENDATIONS

The City should address the problems discussed above by no longer viewing SDPD as an institution that makes Southeast San Diego safer. Rather, SDPD undermines community wellbeing, and inhibits opportunities to thrive, in Southeast San Diego–especially for people of color. To address these issues, provided below are a comprehensive set of equity-centered recommendations that should be adopted. 

## 1. Reinvest Dollars Wasted on Gang Profiling in Equity-Centered Safety Programs

Long-term solutions to improve safety in Southeast San Diego should be grounded in investments that address the root causes of safety risks. Continuing to double-down on law enforcement and incarceration only serves to produce recidivism and disproportionately harm people of color. Rather than annually wasting millions of public dollars on ineffective and racially biased gang profiling, the City should shift public dollars to programs that address fundamental needs of community members in Southeast San Diego, including affordable housing, jobs that pay living wages, workforce development, and food security. For example, research shows that permanent supportive housing and guaranteed basic income programs targeted to historically under-resourced communities can reduce safety risks.^[https://slate.com/news-and-politics/2022/03/how-cash-transfer-programs-prevent-violent-crime.html] This would be far more effective than criminalizing Black and Latinx people in Southeast San Diego. In addition, community-based organizations that serve Southeast San Diego have a deep understanding of the unique challenges community members face. Their experience and expertise as trusted messengers and catalysts for safety should be leveraged by providing them with greater resources to help advance safety. 

## 2. Enact the PrOtect Act

The City should enact the PrOtect (Preventing Over-policing through Equitable Community Treatment) Act.^[https://prismic-io.s3.amazonaws.com/protect/c8a963b8-93e3-482c-abc5-2f5763b1cd82_PrOTECT+Ordinance+%28v.+8.4+web%29.pdf] Developed as a priority of the Coalition for Police Accountability & Transparency, the PrOtect Act aims to prevent racially biased stops and searches. Its key provisions include requiring officers to have probable cause to stop a person (including for people on probation or parole), require probable cause for searches, prohibiting questioning for offenses beyond the offense that gives rise to a stop, holding officers accountable for violations of those restrictions, and publicly reporting data on stops.^[https://d3n8a8pro7vhmx.cloudfront.net/alliancesandiego/pages/204/attachments/original/1617815407/PrOTECT_Factsheet_%282%29.pdf?1617815407]

## 3. Acknowledge Harms Caused by Gang Profiling and Apologize to Southeast San Diego Community Members

The City and SDPD should issue a public apology for harms inflicted upon Southeast community members through racially biased gang profiling. While an apology alone is insufficient to fully address past harms, it is a key reparative stepping stone.^[https://oag.ca.gov/system/files/media/ch16-ca-reparations.pdf 
 https://www.americanprogress.org/article/truth-and-reconciliation] With roots in South African apartheid, there is a lengthy history of governmental apologies for dehumanizing atrocities. For example, in 2023, the Chief of Police of the Seattle Police Department (SPD) provided a formal public apology for SPD’s violent response to protesters demanding justice for police violence.^[https://ny1.com/nyc/all-boroughs/ap-top-news/2023/04/19/panel-seattle-police-should-apologize-for-protest-violence] Along the same line, in 2016, the President of the International Association of Chiefs of Police (IACP)–the largest police management organization in the U.S.–issued a formal apology at the IACP’s annual conference in San Diego. He declared “For our part, the first step is for law enforcement and the IACP to acknowledge and apologize for the action of the past and the role that our profession has played in society’s historical mistreatment of communities of color.”^[https://www.washingtonpost.com/news/true-crime/wp/2016/10/17/head-of-u-s-police-chiefs-apologizes-for-historic-mistreatment-of-minorities/]

## 4. Remove All Gang Units from Southeast San Diego

SDPD’s gang units are notorious for inflicting many of the most harmful actions inflicted on community members in Southeast San Diego. As explained above, this includes not only unwarranted stops that undermine safety and waste public dollars, but also harassment, trauma, uses of force, and other forms of dehumanizing treatment. Thus, that gang unit should be prohibited from patrolling in Southeast San Diego or altogether disbanded. Any issues related to gangs would be better addressed through the City increasing investments in community-based organizations and service providers that specialize in gang intervention and violence prevention.

## 5. Enhance SPDP Accountability

SDPD officers who have harmed community members in Southeast San Diego through racial profiling, harassment, uses of force and other forms of physical, mental, and trauma should be held accountable for their actions.

## 6. Increase SDPD Transparency

SDPD declined to provide stop data sought by Catalyst California through a public records request. The data sought would have enabled a more complete analysis of gang profiling by SDPD. The Department’s lack of transparency masks the extent of gang profiling by SDPD and attendant racially biased harms. SDPD’s secrecy regarding stop data that should be made publicly available erodes public trust. 

## 7. Decriminalize Non-Safety Related Quality-of-Life Issues

While this report focuses on SDPD’s gang profiling activities, many of the stops that give rise to gang profiling are based on minor, non-safety related issues, such as vehicle equipment or administrative violations, or quality-of-life concerns that do not harm others (such as the presence of people who are unhoused). The City should thoroughly examine offenses that give rise to gang profiling and eliminate any local ordinances or municipal violations used for profiling that pose little-to-no safety risks. 


# CONCLUSION

Community safety in Southeast San Diego is undermined by SDPD’s racially biased gang profiling. Data and stories from community members show that Black and Latinx are disproportionately harmed and forced to endure severe physical, mental, and emotional trauma. To add, SDPD harmful and unproductive gang profiling wastes tremendous public resources that should be better invested in equity- and care-centered safety programs that do not rely on law enforcement. The City should address these issues by adopting the comprehensive set recommendations provided above.


# AUTHORS

**Research and Data Analysis**

Elycia Mulholland Graves, Director, Research & Data Analysis, Catalyst California

Jennifer Zhang, Senior Research & Data Analyst, Catalyst California

David Segovia, Research & Data Analyst, Catalyst California

Hillary Khan, Data Architect, Catalyst California

**Authors**

Elycia Mulholland Graves, Director, Research & Data Analysis, Catalyst California

Chauncee Smith, Associate Director, Reimagine Justice & Safety, Catalyst California

Mitchelle Woodson, Legal Director, Pillars of the Community

Jennifer Zhang, Senior Research & Data Analyst, Catalyst California

**Communications**

Ron Simms, Jr., Associate Director, Communications, Catalyst California

Roxanna Reyes, Senior Digital Communications Associate, Catalyst California

**Video Storytelling**

[List videographer, and other Pillars would like to have acknowledged]


# METHODOLOGY

[link to methodology on GitHub]