---
title: "Joint Report with Pillars of the Community on SDPD Gang Enforcement"
subtitle: "DRAFT"
author: "Catalyst California and Pillars of the Community"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
    toc: yes
    toc_depth: 2
    toc_float: yes
---
  
<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)


list.of.packages <- c("usethis", "highcharter", "tidyverse", "sf", "leaflet", "sp", "htmltools", "magrittr", "flextable", "lorem",  "htmlwidgets", "vembedr", "knitr")

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

#### Load libraries ####
library(usethis)
library(highcharter)
library(tidyverse)
library(sf)
library(leaflet)
library(sp)
library(htmltools) # for html escape
library(magrittr)
library(flextable)
library(lorem)
library(htmlwidgets)
library(vembedr)
library(knitr)

#### Add database connection ####
source("W:\\RDA Team\\R\\credentials_source.R")

con <- connect_to_db("rjs_pillars")

## Source the Chart Template

source("report_embeds/chart_functions.R")

## Reknit all embeds so git will recognize file changes
# source("report_embeds/run_embeds.R")

peridot <- "#CEEA01"
orange <- "#F25922"

```

*The San Diego Police Department criminalizes communities of color, especially Black people, in Southeast San Diego by routinely labeling them as gang members. This practice inflicts devastating racially biased harms, undermines community safety, and wastes tremendous public dollars.*  


# ACKNOWLEDGEMENTS

Thank you to the people of Southeast San Diego for sharing their stories. Their stories bring to light the inhumane interactions the Southeast community experiences daily with the police. We are incredibly grateful for their courage and honesty. We urge the City to respond to their stories by putting an end to racially biased policing in San Diego.

This project was completed with support from the Microsoft Justice Reform Initiative. 

```{r}
embed_url("https://youtu.be/cWlR-a1bm5w")

```

# CONTEXT

**The San Diego Police Department ("SDPD")**, and law enforcement agencies throughout California have a long history of unjustly criminalizing people of color through gang profiling. This occurs through officers documenting community members–especially those who are Black and/or Latinx as a member or affiliate of a gang and placing their identities in electronic gang databases.  In 1987, for instance, the Los Angeles Sheriff’s Department (“LASD”) and the Law Enforcement Communication Network (NALECOM) – a rapid interstate law enforcement communications system^[https://www.ojp.gov/ncjrs/virtual-library/abstracts/national-law-enforcement-communication-network-nalecom-users] – jointly developed a database used to collect, retain, and analyze information about alleged gang members.^[See Tracked & Trapped at 2.] Over the next decade, law enforcement agencies throughout California refined and expanded gang databases.^[Id. at 3.]

In 1997, a company called CSRA International, Inc. (“CSRA”) used funding from the California Department of Justice to develop the most prominent system – “CalGang” (also known as “GangNet”). CalGang allows law enforcement throughout California to collect, store, and share personal information about people suspected – but not convicted – of being involved in criminal activity.^[State Auditor at 2.] People added to CalGang are documented as either a gang member or gang affiliated.^[Id. at 11.] In addition to in-state law enforcement agencies, CalGang is used by federal law enforcement, other states, and Canada.^[Id. at 9.] Reports indicate that in 2022, at least 106 law enforcement agencies used CalGang, including the San Diego Police Department (SDPD). Amongst those agencies, SDPD had the third highest number of records added to the system (only the LASD  and San Bernardino County Sheriff’s Department had more).      

```{r calgang-table, child= 'report_embeds/report_calgangstats_table.Rmd', echo=FALSE}
```

**The process through which people of color are gang profiled is notoriously problematic.**   Many people have been added to gang databases without prior knowledge and without being arrested for a crime.  For example, officers often conduct pretextual stops of Black and Latinx people by detaining them for relatively minor issues–such as a broken taillight or jaywalking–and interrogating them in aim of discovering evidence or a more serious crime.  During these encounters, officers document personal information about the community member on a Field Interview (FI) card.  The information collected may include, but not be limited to, the person’s name, date of birth, race, gender, nicknames, their family members and friends, their place of residence, the community where they live, photos, tattoos, clothing color, vehicles, and suspected gang affiliation.^[Tracked & Trapped at 4-5. State Auditor at 11.] These interactions and resulting implications impose devastating, long-lasting harms on the community members involved.

```{r}
embed_url("https://www.youtube.com/watch?v=FTqt90YJ_V4")

```

<div style="background-color:#F25922">
## The devastating impact of criminalizing people
Criminalization occurs when people are turned into criminals by making their alleged activities illegal, even if the activities are ordinary or not harmful to others.^[https://www.oed.com/search/dictionary/?scope=Entries&q=criminalization+ ; https://www.merriam-webster.com/dictionary/criminalize] SDPD criminalizes community members from Southeast San Diego by documenting them as gang members. This inflicts devastating harms, direct and indirect, on Southeast community members and their families–including physical, mental, and emotional trauma.  

In addition, being documented as a gang member can have devastating impacts on a person being charged with an alleged crime because of sentencing enhancements for gang membership (or “gang enhancements”). Specifically, under the STEP Act, the application of a gang enhancement adds an additional two years to life imprisonment to a person’s sentence.  The added time runs consecutive (rather than concurrent) to the punishment for an underlying felony.^[Yoshino at 119.] The potential for such added harm often has devastating impacts on people facing relatively low charges because the threat of far more lengthy imprisonment impacts plea bargaining, charging decisions, juror perspectives, among other things.   
</div>


# BRIEF METHODOLOGY

This report evaluates SDPD gang profiling in the City of San Diego by analyzing data collected by SDPD pursuant to the Racial and Identity Profiling Act (“RIPA”) of 2015.  For each stop made by law enforcement, RIPA data includes key information that can be assessed for evidence of racial or identify profiling—including characteristics about the person stopped (e.g., race, gender, and age), the stop location, the length of the stop, the reason for the stop, and the result of the stop. 

Using RIPA data, this report examines gang profiling by using the officer assignment field (under “gang enforcement”) which we find underrepresents total gang enforcement activity being done by SDPD, and stop result filed (under “field interview card”). 


# FINDINGS

In sum, the data for SDPD clearly show racial bias towards people who officers perceive as Black and/or Latinx, and that these groups are disproportionately impacted by gang enforcement activity. Regions with higher concentrations of  Black and Latinx residents, such as Southeast San Diego, have higher stop rates and are subject to greater gang enforcement activity. Furthermore, comparing stop rates to the San Diego population shows that SDPD stops people perceived as Black at rates much higher than their proportion of the population. In addition to racial bias, the data show that SDPD gang profiling activities often undermine safety by inflicting tremendous harm upon Southeast San Diego community members while rarely resulting in the discovery of criminal evidence or an arrest. . Community member stories add more poignancy to the impact of SDPD’s racially biased activities by showing how Black and Latinx people have personally been harmed repeatedly by officers.

## 1. SDPD targets communities of color, particularly Black and young people in Southeast San Diego, with gang profiling

In Southeast San Diego, SDPD’s practice of racial profiling people for gang activity is far too common and affects the daily lives of community members. Lived experiences reveal the long-lasting harm, trauma, and distrust caused by these practices.

```{r}
embed_url("https://www.youtube.com/watch?v=arZekGIgIrw")

```

### A. Data show that stops made by the SDPD are racially biased

An analysis of all officer-initiated stops–as opposed to stops based on service calls–shows that officers disproportionately stop people of color, particularly people officers perceive as Black or Native Hawaiian or Pacific Islander (NHPI). More specifically, in 2022, for every one-thousand Black people in San Diego, officers stopped nearly 216 people they perceived as Black. In comparison, the average stop rate in San Diego in 2022 was 64 people for every one-thousand people in the city. This means that SDPD was more than three times as likely to stop a Black person than city residents at large. And, while Black people comprise 5.7 percent of the total population, they accounted for 19.2 percent of officer-initiated stops.

```{r bubble-chart, child= 'report_embeds/report_stoprates_race_person_bubble.Rmd', echo=FALSE}
```

A common assumption is that officer-initiated stops advance safety. However, data show that officer-initiated stops are often driven by racial bias rather than safety risks. Stops that result in no enforcement action or a relatively minor result (such as a warning) indicate relatively low safety risks. When people of color are disproportionately subjected to stops that pose little to no safety risks, it indicates racial bias. These stops, known as pretextual stops, are often based on a racially biased and incorrect “hunch” that a person of color was connected to illegal activity. All too often, pretextual stops are used to expose people of color to unnecessary contact with armed law enforcement.

In San Diego, Black and Latinx people are disproportionately subjected to officer-initiated stops that result in no action or a warning. Specifically, in 2022, nearly 1 in 4 stops (25.3%) of people officers perceived as Black resulted in a warning and over 1 in 10 (12.7%) resulted in no action. Combined these stops affected an estimated 6,416 Black people in San Diego. People officers perceived as Latinx have the second highest rate of stops resulting in no action (11.5%) or a warning (26.8%). These racially biased stops affected an estimated total of 11,563 Latinx people.

```{r bubble-chart, child= 'report_embeds/report_stoprates_result_race_person_stacked.Rmd', echo=FALSE}
```

### B. Data show that SDPD officers assigned to gang enforcement profile Black and Latinx people   

Stories from community members reveal that SDPD’s gang enforcement officers target and harass Black and Latinx people and the Southeast community and label them as gang members. And many times officers target young people engaged in everyday activities with no indication of safety risks.

```{r}
embed_url("https://www.youtube.com/watch?v=Dd-e07UT6-Y")

```

Data  show that officers assigned to gang enforcement at the time of the stop target Latinx and Black people. Specifically, amongst all people stopped by officers assigned to gang enforcement, 40.1% were perceived as Latinx and 32.4% were perceived as Black. Compared to their share of the San Diego population, Black and Latinx people were respectively 5.7 and 1.3 times overrepresented in people stopped by officers assigned to gang enforcement. 

Black people of every age group in San Diego are disproportionately impacted by these stops. For instance, while Black people ages 45-54 comprise just 5.2% of the San Diego population in that age group, they made up 41.5% of people stopped by officers assigned to gang enforcement. This translates into Black people ages 45-54 being nearly 8 times overrepresented in people stopped by officers assigned to gang enforcement. Latinx people between the ages of 18-44 are over 1.3 times overrepresented in people stopped by officers assigned to gang enforcement compared to their share of the San Diego people. These trends indicate the alarming rate of racial profiling used by the Gang Suppression Unit. However, the counts included in this data are likely underinclusive of the actual number of stops done by the Gang Suppression Unit.

```{r bubble-chart, child= 'report_embeds/report_stoprates_race_gang_person_bubble.Rmd'}
```


<div style="background-color:#F25922">
### SDPD data are underinclusive and mask the extent of racially biased, gang profiling in San Diego

Our research suggests that publicly available SDPD data on gang documentation are underinclusive and, in turn, mask the scope of racial biases in SDPD gang enforcement. This is in part due to a lack of clarity (and inconsistencies) between the officer assignment categories used under RIPA regulations, SDPD’s list of units, and the functional tasks SDPD’s officers are responsible for irrespective of their assignments. 

More specifically, RIPA regulations require that the “type of assignment of officer” be reported for each stop.^[11 Cal. Code Regs. § 999.226(a)(20)] There are eleven assignment types, including “gang enforcement.” Catalyst California’s research indicates that SDPD has approximately 39 departmental units to which officers are assigned.  This includes a “gang unit” for personnel that, among other things, identify and investigate gang members and gang activity, maintain SDPD gang information files and liaise with agencies across the nation, and assist patrol with gang information.^[https://www.sandiego.gov/sites/default/files/legacy/police/pdf/unitlist.pdf] This latter function–assisting patrol with gang information–indicates that the gang unit collaborates with other parts of SDPD that conduct patrol activities (such as traffic-related units) on matters related to alleged gang activity that may not qualify as “gang enforcement” under RIPA’s regulation on officer assignment.

To account for this, Catalyst California submitted a public records request to SDPD to determine whether there were additional stops related to gang enforcement that may not have been reported under RIPA’s gang enforcement assignment due to the lack of clarity between RIPA and SDPD definitions and personnel functions. Both requests were denied.
^[insert brief description of SPDP’s responses] SDPD’s lack of transparency  with RIPA data masks the extent of racial biases in SDPD’s gang profiling and, in turn, attendant harms that are disproportionately experienced by Black and Latinx community members in Southeast San Diego.   

Because SDPD refused to provide data specific to its Gang Suppression Unit, the data was analyzed based on stops that result in a field interview card as a proxy for identifying  gang suppression activity by SDPD. This data, combined with stories from impacted community members, further reveal the extent of extremely harmful SDPD’s gang profiling practices
</div>

```{r}
embed_url("https://www.youtube.com/watch?v=sD5libtgXfI")

```

### C. Beyond officers assigned to gang enforcement, an analysis of stops that result in field interview cards shows that SDPD targets Black and Latinx people and labels them as gang members 

Community members’ experiences show how SDPD’s gang profiling activity subjects them to traumatic encounters that go far beyond what SDPD reports. These encounters often include field interview cards that are used to enter community members into CalGang. 

```{r}
embed_url("https://www.youtube.com/watch?v=oOmzLOEOFGM")

```

Data shows that SDPD stops people they perceive as Black for field interview cards at the highest rate compared to all other racial groups. While Black people account for 5.6% of San Diego’s total population, White people make up 42% of the population. The rate at which SDPD stops people they perceive as Black for field interview cards is 5 times more than the rate at which they stop people perceived as White. 

Communities most impacted by SDPD gang profiling report that the department targets youth of color. The data confirms this trend, and shows alarming racial disparities across other age groups. Young adults ages 18-24 and minors 17 years old and younger are more likely to be stopped and have a field interview card completed than all other age groups. And, across age ranges, SDPD stops and completes field interview cards for people they perceive as Black or Latinx at higher rates compared to other racial groups.

```{r bubble-chart, child= 'report_embeds/report_stoprates_ficard_race_person_bubble.Rmd'}
```

SDPD’s racial targeting for stops that result in a field interview card primarily takes place in regions nearSouth and Southeast San Diego, where more people of color live. Analyzing the data by police division shows these geographic disparities., SDPD makes the most amount of stops that result in a field interview card in the Central, Southeastern, and Mid-City police divisions.  In 2022, SDPD made 1,100 total stops with a field interview card in the Southeastern division (e.g., Lincoln Park, Skyline, etc.) compared to only 171 total stops resulting in a field interview card in the Northwestern division. San Diego’s Whiter and more affluent communities (e.g., Carmel Valley, and Del Mar Heights) are in the Northwestern division . Black people make up less than 1% of the population in the Northwestern division are Black, but 14.9% of the population in Southeast San Diego.

Analysis by race shows that the people SDPD disproportionately stops for field interview cards are people of color, especially people perceived as Black and Latinx. In the Southeastern division, 1 in 5 people (19.4%) SDPD stopped and perceived as Black have field interview cards completed, compared to less than 1 in 10 people perceived as White.

```{r dot-density, child= 'report_embeds/report_stoprates_div_ficard_race_person_density_map.Rmd'}
```

In addition, the data indicate that SDPD’s practice of disproportionately stopping people of color and subjecting them to field interviews (which often amount to interrogation and harassment) is rooted in racial profiling. Specifically, nearly 1 in 4 stops that resulted in a field interview card originated based on alleged traffic violations. Analyzing those stops by race shows that SDPD used traffic violations to stop people they perceived as people of color for field interview cards at more than twice the rate of people perceived as White. To add, while these stops were allegedly made in response to traffic safety risks, they concluded in results that indicate otherwise because no citations were issued. Rather, the officer only completed field interview cards. This indicates pretext and the use of racial profiling to harass people of color for gang activity.

```{r stacked-chart, child= 'report_embeds/report_stoprates_ficard_reason_race_bubble.Rmd'}
```

## 2. Such racially biased stops harm the Southeast community and undermine community safety  

SDPD’s gang profiling activity unnecessarily subjects community members to harmful contact with armed officers and undermines community safety. Stories from community members show that they are often harassed, interrogated, physically searched, and subjected to other degrading and dehumanizing treatment by officers. This results in community members feeling unsafe from the institution that is supposed to protect them from harm. 

```{r}
embed_url("https://www.youtube.com/watch?v=e1KxqPIKjeo")

```

### A. Searches conducted during gang profiling are largely ineffective and racially biased 
During stops, officers often search people who they think may possess evidence of a crime. Data on police stops repeatedly show that officer searches are highly ineffective in rooting out crime and are racially biased. Specifically, amongst all people searched in 2022, officers failed to find evidence of a crime 70 percent of the time. In other words, over 18,000 people were subjected to a police search by SDPD in 2022 that was proven unnecessary.

```{r donut-chart, child= 'report_embeds/report_hitrates_race_person_donut.Rmd'} 
```

To add, searches conducted during stops that result in a field interview card were even less effective. Officers failed to find evidence of a crime in approximately 72 percent of stops that resulted in a field interview. Considering race and ethnicity in stops involving a search and resulting in a field interview, no contraband was found in 84 percent of stops of people perceived as NHPI. This rate drops to 65 percent during field interview card stops of people perceived as White, showing that searches for contraband were less effective when applied to people of color. These racially biased and unnecessary searches and contact with officers affected the highest numbers of Latinx and Black people the most. In 2022, officers subjected 1,705 people perceived as Latinx and 1,366 people perceived as Black to these searches.

```{r stacked-chart, child= 'report_embeds/report_hitrates_race_ficard_person_bubblepop.Rmd'}
```

### B. Southeast community members have had their wellbeing and safety harmed by gang enforcement stops and searches

While these SDPD practices occur throughout San Diego, data and stories from the Southeast community show the prevalence of these searches in Southeast San Diego. Community members describe how physical interactions with police creates emotional trauma and physical harm. 

```{r}
embed_url("https://www.youtube.com/watch?v=KzE0t3U4Sk0")

```

Division-level analysis of searches done during field interview card stops further demonstrates the extent of this harmful practice and how it targets people of color in Southeast San Diego. In all of San Diego, there were 2,936 people who were searched, field interviewed, and found to have no contraband. Of these people, around 66 percent were in the Southeastern, Central, and Mid-City divisions of San Diego, areas that have a higher concentration of people of color compared to the Northern regions.  

The Southeast Division has the highest number of people searched during field interview card stops and the highest number of people found with no contraband. SDPD officers found no contraband in 77 percent of searches done during stops where a field interview card was completed in the Southeastern Division. Black and Latinx people in the Southeast were disproportionately targeted by these searches. People perceived as Black and Latinx accounted for over 89 percent of these unsuccessful searches while only making up 65 percent of the population in this division. These disparities existed in other divisions as well. While Black people represented only seven percent of the population in the Central division, they comprised 35 percent of people searched with no contraband or evidence. And while they represented the second highest searched group in the Central Division, only second to Latinx individuals, officers failed to find contraband or evidence on nearly 81 percent of people they perceived as Black who they searched and field interviewed.

```{r dot-map, child= 'report_embeds/report_hitrates_div_race_ficard_person_density_map.Rmd'}
```

## 3. These racially biased activities waste public resources

Like most police departments in California, a significant portion of public dollars goes to SDPD. The narrative pushed by police departments is that a huge amount of funding is needed to keep communities safe. However, data shows that police spend their time on activities that are not preventing crime or keeping the public safe. SDPD is no exception to this.


```{r}
embed_url("https://www.youtube.com/watch?v=W_j4T1uOQhY")

```

### A. Based on available data, SDPD spends a disproportionate amount of time on profiling people of color for gang activity

When analyzing how much time SDPD spends on stops resulting in a field interview card, nearly six percent of all hours spent on officer-initiated stops are spent on stops that result in a field interview card.  

Within stops that result in a field interview card, SDPD disproportionately spends their time stopping people perceived as Latinx and Black. More than half of the hours spent on stops that result in a field interview card are spent stopping people perceived as Latinx or Black, even though Latinx and Black people make up only 35.7 percent of the total population in San Diego.

Black community members throughout San Diego are overwhelmingly impacted by this racially biased use of public resources. SDPD spends more time conducting field interview cards for Black people relative to their population in every police division. In Southeast San Diego, officers spent three times more hours on field interview cards of Black people compared to their relative share of the population in Southeast San Diego.  

```{r item-chart, child= 'report_embeds/report_timespent_div_ficard_race_person_group.Rmd'}
```

Analysis of hours spent by SDPD officers assigned to gang enforcement also show that most of their time is not being spent on activities that prevent crime. Offices assigned to gang enforcement spent nearly 4 in 10 hours on stops resulting only in a field interview card, a warning, or no action. Stops only resulting in a citation for an infraction took up an additional 8 out of 100 patrol hours. Comparatively, arrests with warrants only comprised 9 in 100 patrol hours. 

```{r item-chart, child= 'report_embeds/report_timespent_ficard_gang_stop_item.Rmd'}
```

### B. This time translates to wasted public dollars.

`r lorem::ipsum(paragraphs = 1,sentences=NULL, avg_words_per_sentence=10)`


# RECOMMENDATIONS

1. Alternative investments that actually increase safety
2. Enact the Protect Act
3. Others...


# AUTHORS

**Research and Data Analysis**

Elycia Mulholland Graves, Director, Research & Data Analysis, Catalyst California

Jennifer Zhang, Senior Research & Data Analyst, Catalyst California

David Segovia, Research & Data Analyst, Catalyst California

Hillary Khan, Data Architect, Catalyst California

**Authors**

Elycia Mulholland Graves, Director, Research & Data Analysis, Catalyst California

Chauncee Smith, Associate Director, Reimagine Justice & Safety, Catalyst California

Mitchelle Woodson, Legal Director, Pillars of the Community

Jennifer Zhang, Senior Research & Data Analyst, Catalyst California

**Communications**

Ron Simms, Jr., Associate Director, Communications, Catalyst California

Roxanna Reyes, Senior Digital Communications Associate, Catalyst California

**Video Storytelling**

[List videographer, and other Pillars would like to have acknowledged]


# METHODOLOGY

[link to methodology on GitHub]