---
title: "Joint Report on SDPD's Gang Enforcement Practices"
subtitle: "DRAFT"
author: "Catalyst California and Pillars of the Community"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
    toc: yes
    toc_depth: 2
    toc_float: yes
---
  
<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)


list.of.packages <- c("usethis", "highcharter", "tidyverse", "sf", "leaflet", "sp", "htmltools", "magrittr", "flextable", "htmlwidgets")

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

#### Load libraries ####
library(usethis)
library(highcharter)
library(tidyverse)
library(sf)
library(leaflet)
library(sp)
library(htmltools) # for html escape
library(magrittr)
library(flextable)
library(lorem)
library(htmlwidgets)
library(vembedr)

#### Add database connection ####
source("W:\\RDA Team\\R\\credentials_source.R")

con <- connect_to_db("rjs_pillars")

## Source the Chart Template

source("report_embeds/chart_functions.R")

## Reknit all embeds so git will recognize file changes
# source("report_embeds/run_embeds.R")

peridot <- "#CEEA01"
orange <- "#F25922"

```

*The San Diego Police Department criminalizes communities of color, especially Black people, in Southeast San Diego by routinely labeling them as gang members. This practice inflicts devastating racially biased harms, undermines community safety, and wastes tremendous public dollars.*  

[this report section–add info about this work, partners, etc.]

# ACKNOWLEDGEMENTS

[Acknowledgements–Acknowledge and thank people of Southeast San Diego for sharing their stories. Include acknowledgement of funding from Microsoft/Urban]

<html>
<iframe src="https://drive.google.com/file/d/1gn5c8mytLXBpWlYKphRTI_jZyxUYpPvh/preview" width="640" height="480" allow="autoplay"></iframe>
</html>

```{r,include=FALSE}
embed_url("https://www.youtube.com/watch?v=q1gicHIVxV8")

```

# CONTEXT

Law enforcement agencies throughout California have a long history of unjustly criminalizing people of color—especially  those who are Black and/or Latinx—by documenting them as a member or affiliate of a gang and placing their identities in electronic gang databases.  In 1987, for instance, the Los Angeles Sheriff’s Department and the Law Enforcement Communication Network (NALECOM) – a rapid interstate law enforcement communications system^[https://www.ojp.gov/ncjrs/virtual-library/abstracts/national-law-enforcement-communication-network-nalecom-users] – jointly developed a database used to collect, retain, and analyze information about alleged gang members.^[See Tracked & Trapped at 2.] Over the next decade, law enforcement agencies throughout California refined and expanded gang databases.^[Id. at 3.]

In 1997, a company called CSRA International, Inc. (“CSRA”) used funding from the California Department of Justice to develop the most prominent system – “CalGang” (also known as “GangNet”).  CalGang allows law enforcement throughout California to collect, store, and share personal information about people suspected – but not convicted – of being involved in criminal activity.^[State Auditor at 2.] People added to CalGang are documented as either a gang member or gang affiliated.^[Id. at 11.] In addition to in-state law enforcement agencies, CalGang is used by federal law enforcement, other states, and Canada.^[Id. at 9.] Reports indicate that in 2022, at least 106 law enforcement agencies used CalGang, including the San Diego Police Department (SDPD). Out of those agencies, SDPD had the third highest number of records added to the system with only San Bernardino and Los Angeles Counties’ Sheriff’s Departments exceeding their numbers.      

```{r calgang-table, child= 'report_embeds/report_calgangstats_table.Rmd', echo=FALSE}
```

Research shows that many people have been added to gang databases without prior knowledge and without being arrested for any crime.  Officers often conduct pretextual stops of Black and Latinx people, interrogate them, and document their personal information on a Field Interview (FI) card.  The information collected may include, but not be limited to, the person’s name, date of birth, race, gender, nicknames, their family members and friends, their place of residence, the community where they live, photos, tattoos, clothing color, vehicles, and suspected gang affiliation.^[Tracked & Trapped at 4-5. State Auditor at 11.]

<p style="background-color:#F25922;font-size:14pt">Criminalization occurs when people are turned into criminals by making their alleged activities illegal, even if the activities are ordinary or not harmful to others.^[https://www.oed.com/search/dictionary/?scope=Entries&q=criminalization+ ; https://www.merriam-webster.com/dictionary/criminalize] SDPD criminalizes community members from Southeast San Diego by documenting them as gang members. This inflicts devastating harms, direct and indirect, on Southeast community members and their families.</p>

<html>
<iframe src="https://drive.google.com/file/d/1z2fxt1zI10t0MbE4ekskMV-cRlGjjVgT/preview" width="640" height="480" allow="autoplay"></iframe>
</html>

# FINDINGS

[brief description of methodology]

## 1. SDPD targets communities of color, particularly Black and young people in Southeast San Diego, with this harmful practice

In Southeast San Diego, SDPD’s practice of racial profiling people for gang activity is far too common and affects the daily lives of many community members. Both lived experiences and police stop data reveal racial disparities and harms caused by these practices.

<html>
<iframe src="https://drive.google.com/file/d/1J9_1B30d145x20qVFcjMo6B1-R6EGE11/preview" width="640" height="480" allow="autoplay"></iframe>
</html>

### A. Data show that stops made by the SDPD are racially biased.

An analysis of all stops voluntarily initiated by officers, referred to as officer-initiated stops, shows that officers disproportionately stop people of color, particularly people officers perceive as Black or Native Hawaiian or Pacific Islander (NHPI). In 2022, for every thousand Black people in San Diego, officers stopped nearly 216 people they perceived as Black. In comparison, the average stop rate in San Diego in 2022 was 64 people for every thousand people in the city. Meaning, officers are more than three times as likely to stop a Black person in San Diego than city residents at large. And while Black people comprise only 5.7 percent of the total population in San Diego, they accounted for 19.2 percent of officer-initiated stops.

```{r bubble-chart, child= 'report_embeds/report_stoprates_race_person_bubble.Rmd', echo=FALSE}
```

A common assumption is that officer-initiated stops advance safety. However, data show that officer-initiated stops are often driven by racial bias rather than safety risks. Stops that result in no enforcement action or a relatively minor result (such as a warning) indicate  relatively low safety risks.  When people of color are disproportionately subjected to stops that result in no action or a warning (i.e., stopped for activities that pose little to no safety risk), it indicates racial bias.  These stops, known as pretextual stops, show evidence that an officer initiated a stop based on a “hunch” that a person was connected to illegal activity but found no serious safety risk once the stop was made. Pretextual stops are consistently used against people of color, and expose them to unnecessary contact with armed law enforcement. 

In San Diego, Black and Latinx people are disproportionately subjected to officer-initiated stops that result in no action or a warning. Specifically, in 2022, nearly 1 in 4 stops (25.3%) of people officers perceived as Black resulted in a warning and over 1 in 10 (12.7%) resulted in no action. Combined these stops affected an estimated 6,416 Black people in San Diego. People officers perceived as Latinx have the second highest rate of stops resulting in no action (11.5%) or a warning (26.8%). These racially biased stops affected an estimated total of 11,563 Latinx people.

```{r bubble-chart, child= 'report_embeds/report_stoprates_result_race_person_stacked.Rmd', echo=FALSE}
```

### B. Available data show that SDPD’s gang enforcement assignment profiles Black and Latinx people for gang activity.   

Stories from community members indicate the practices of SDPD’s Gang Suppression Unit are even more troubling. Gang enforcement officers target and harass Black and Latinx people and the Southeast community and label them as gang members.

<html>
<iframe src="https://drive.google.com/file/d/16hHoLrYxwmL8ZiVOs8DUfy6t4gMSv-R0/preview" width="640" height="480" allow="autoplay"></iframe>
</html>

Analysis shows that officers assigned to gang enforcement at the time of the stop target Latinx and Black people. Specifically, amongst all people stopped by officers assigned to gang enforcement, 40.1% are people officers perceived as Latinx and 32.4% are people perceived as Black. Compared to their share of the San Diego population, Black and Latinx people are 5.7 and 1.3 times overrepresented in people stopped by officers assigned to gang enforcement.  

Black people of every age group in San Diego are disproportionately impacted by these stops. For instance, while Black people ages 45-54 comprise just 5.2% of the San Diego population in that age group, they made up 41.5% of people stopped by officers assigned to gang enforcement. This translates into Black people ages 45-54 being nearly 8 times overrepresented in people stopped by officers assigned to gang enforcement. Latinx people between the ages of 18-44 are over 1.3 times overrepresented in people stopped by officers assigned to gang enforcement compared to their share of the San Diego people. These trends indicate the alarming rate of racial profiling used by the Gang Suppression Unit. However, the counts included in this data are likely underinclusive of the actual number of stops done by the Gang Suppression Unit

```{r bubble-chart, child= 'report_embeds/report_stoprates_race_gang_person_bubble.Rmd'}
```


<div style="background-color:#F25922">
### SDPD data are underinclusive and mask the extent of racially biased, gang profiling in San Diego.

Our research suggests that SDPD data on gang documentation are underinclusive and, in turn, mask scope of racial biases. This is in part due to a lack of clarity (and inconsistencies) between the officer assignment categories used under RIPA regulations, SDPD’s list of units, and the functional tasks SDPD’s officers are responsible for irrespective of their assignments. 

More specifically, RIPA regulations require that the “type of assignment of officer” be reported for each stop.^[11 Cal. Code Regs. § 999.226(a)(20)] It includes eleven options, including “gang enforcement.” Our research indicates that SDPD has approximately 39 departmental units to which officers are assigned. This includes a “gang unit” for personnel that, among other things, identify and investigate gang members and gang activity, maintain SDPD gang information files and liaise with agencies across the nation, and assist patrol with gang information.^[https://www.sandiego.gov/sites/default/files/legacy/police/pdf/unitlist.pdf] This latter function, that is, assisting patrol with gang information, indicates that the gang unit collaborates with other parts of SDPD that conduct patrol activities (such as traffic-related units) on matters related to alleged gang activity that may not qualify as “gang enforcement” under RIPA’s regulation on officer assignment.

To account for this, the authors of this report submitted two public records to SDPD to determine whether there were additional stops related to gang enforcement that may not have been reported under RIPA’s gang enforcement assignment due to the lack of clarity between RIPA and SDPD definitions and personnel functions.  Both requests were denied.^[insert brief description of SPDP’s responses]  SDPD’s failure to be transparent with RIPA data on stops masks the extent of racial biases in SDPD’s gang documentation activities and, in turn, attendant harms that are disproportionately experienced by Black and Latinx community members in Southeast San Diego.   

Because SDPD refused to provide data specific to its Gang Suppression Unit, we use stops that result in a field interview card as a proxy for identifying  gang suppression activity by SDPD. We combine this data with stories from impacted community members to show the extent of this harmful practice.
</div>

<html>
<iframe src="https://drive.google.com/file/d/1-SVjkbS3zQtwH_ikKBi2SIF5vOQIIUUf/preview" width="640" height="480" allow="autoplay"></iframe>
</html>

### C. Beyond officers assigned to gang enforcement, an analysis of stops that result in field interview cards also show that SPDP targets Black and Latinx people and labels them as gang members. 

Community members’ experiences show how SDPD’s gang profiling activity subjects them to traumatic encounters that go far beyond  what SDPD reports. These encounters often include field interview cards that are used to enter community members into California’s Cal Gang database. 

<html>
<iframe src="https://drive.google.com/file/d/1LPkK-TUHj0WQ7pHk6IcoEGlRKizMkvB6/preview" width="640" height="480" allow="autoplay"></iframe>
</html>

Data shows that SDPD stops people they perceive as Black for field interview cards at the highest rate compared to all other racial groups. Despite Black people making up only 5.6% of San Diego’s total population, the rate at which SDPD stops people they perceive as Black for field interview cards is 5 times more than the rate at which they stop people perceived as White. Meanwhile, White people make up 42% of the city population.  

Communities most impacted by police gang enforcement activity report that SDPD targets youth of color for gang activity. The data confirms this trend and shows alarming racial disparities across  other age groups. Young adults ages 18-24 and minors 17 years old and younger are more likely to be stopped and have a field interview card completed than all other age groups.  And, SDPD stops and completes field interview cards for people they perceive as Black or Latinx at higher rates compared to other racial groups.

```{r bubble-chart, child= 'report_embeds/report_stoprates_ficard_race_person_bubble.Rmd'}
```

SDPD’s racial targeting for stops that result in a field interview card primarily takes place in South and Southeast San Diego, where more people of color live. Across all SDPD police divisions, SDPD makes the most amount of stops that result in a field interview card in the Central, Southeastern, and Mid-City divisions.  In 2022, SDPD made 1,100 total stops with a field interview card in the Southeastern division compared to only 171 total stops resulting in a field interview card in the Northwestern division. San Diego’s Whiter and more affluent communities are in the Northwestern division.

Analysis by race shows that the people SDPD disproportionately stops for field interview cards are people of color, especially people perceived as Black and Latinx. In the Southeastern division, 1 in 5 people (19.4%) SDPD stopped and perceived as Black have field interview cards completed, compared to less than 1 in 10 people perceived as White.

```{r dot-density, child= 'report_embeds/report_stoprates_div_ficard_race_person_density_map.Rmd'}
```

Despite the high rate of people of color stopped with field interview cards completed, the data shows that SDPD makes these stops based on pretext and racial profiling. Nearly 1 in 4 stops that result in a field interview card are originally made for traffic violations. Analyzing those traffic stops by race show that SDPD disproportionately stops people of color based on alleged traffic violations to complete field interview cards (a proxy for gang profiling) . People  perceived as people of color are stopped for a traffic violation and have field interview card completed at more than twice the rate of people who are perceived as White by SDPD. While the officer reportedly made these initial stops in response to a traffic safety risk, they concluded in no result indicative of an actual traffic violation , such as a citation. Instead, the officer only completed a field interview card, indicating pretext and the use of racial profiling to harass people of color for gang activity.

```{r stacked-chart, child= 'report_embeds/report_stoprates_ficard_reason_race_bubble.Rmd'}
```

## 2. Such racially biased stops harm the Southeast community and undermine community safety.  

SDPD’s gang profiling activity subjects community members to unnecessary contact with armed officers and undermines the public and individual safety of the community. Stories from community members show that they are often harassed, interrogated, physically searched, and subjected to other degrading and dehumanizing treatment by officers. This results in numerous harms, including feeling a lack of safety from the institution chartered to protect them. In addition, being a documented as a gang member can have devastating impacts on a person being charged with an alleged crime because of sentencing enhancements for gang membership (or “gang enhancements”). Specifically, under the STEP Act, the application of a gang enhancement adds an additional two years to life imprisonment to a person’s sentence. The added time runs consecutive (rather than concurrent) to the punishment for the underlying felony.^[Yoshino at 119.] The potential for such added harm often has devastating impacts on people facing relatively low charges because the threat of far more lengthy imprisonment impacts plea bargaining, charging decisions, juror perspectives, among other things.

<html>
<iframe src="https://drive.google.com/file/d/1k8JKn6R22oKArmrx5WNl9A8kw_T8x0g4/preview" width="640" height="480" allow="autoplay"></iframe>
</html>

### A. When searches are conducted as a part of gang profiling, they are largely ineffective and biased.  

During stops, officers often search people if they think a person may possess evidence of a crime. Data on police stops repeatedly show that officer searches are highly ineffective in rooting out crime and are racially biased. Amongst all people searched in 2022, officers failed to find evidence of a crime 70 percent of the time. In other words, over 18,000 people were subjected to a police search by SDPD in 2022 that was proven unnecessary.

```{r donut-chart, child= 'report_embeds/report_hitrates_race_person_donut.Rmd'} 
```

This is also the case for searches done during field interview card stops. When officers conduct searches during these stops, they show greater evidence of racial profiling. Officers failed to find evidence of a crime approximately 72 percent of the time when conducting a search during a stops that resulted in a field interview card. In 84 percent of searches done during stops where a field interview card was completed of individuals perceived as NHPI, SDPD officers found no contraband or evidence. This rate drops to 65 percent during field interview card stops of people perceived as White, showing that searches for contraband are less effective when applied to people of color. These unnecessary searches and contact with officers affected the highest numbers of Latinx and Black people. In 2022, officers subjected 1,268 people perceived as Latinx and 995 people perceived as Black to these searches.

```{r stacked-chart, child= 'report_embeds/report_hitrates_race_ficard_person_bubblepop.Rmd'}
```

### B. Southeast community members have had their wellbeing and safety harmed by gang enforcement stops.

While these SDPD practices occur throughout San Diego, data and stories from the Southeast community show the prevalence of these searches in Southeast San Diego. Community members describe how these searches are overwhelmingly prevalent in their neighborhoods and expose them to contact with police that unduly creates trauma and harm for the community. 

<html>
<iframe src="https://drive.google.com/file/d/1SzCa50E3FC0AKbSZFi6l89iO-DtWE3gt/preview" width="640" height="480" allow="autoplay"></iframe>
</html>

Analysis of searches done during field interview card stops further demonstrates the extent of this harmful practice and how it targets people of color in San Diego, and particularly, the Southeast community. In all of San Diego, there were 2,936 people who were searched, field interviewed, and found to have no contraband. Of these people, around 66 percent were in the Southeastern, Central, and Mid-City divisions of San Diego, areas that have a higher concentration of people of color in San Diego compared to the Northern regions of San Diego.  

The Southeast Division has the highest number of people searched during field interview card stops and the highest number of people found with no contraband or evidence. In 77 percent of searches done during stops where a field interview card was completed in the Southeastern Division for all individuals, SDPD officers found no contraband or evidence. Black and Latinx people in the Southeast are disproportionately targeted by these searches. People perceived as Black and Latinx accounted for over 89 percent of these unsuccessful searches while only making up 65 percent of the population in this division. These disparities exist in other divisions as well. While Black people represent only seven percent of the population in the Central division, they comprised 35 percent of people searched with no contraband or evidence. And while they represent the second highest searched group in the Central Division, only second to Latinx individuals, officers failed to find contraband or evidence on nearly 81 percent of people they perceived as Black who they searched and field interviewed.

```{r dot-map, child= 'report_embeds/report_hitrates_div_race_ficard_person_density_map.Rmd'}
```

## 3. These racially biased activities waste public resources.

Like most police departments in California, a significant portion of public dollars goes to SDPD. The narrative pushed by police departments is that a huge amount of funding is needed to keep communities safe. However, data shows that police spend their time on activities that are not preventing crime or keeping the public safe. SDPD is no exception to this.

<html>
<iframe src="https://drive.google.com/file/d/1xTibKteCXov0n5SeDy5PibCJmgCzK2Zs/preview" width="640" height="480" allow="autoplay"></iframe>
</html>

### A. Based on available data, SDPD spends a disproportionate amount of time on profiling people of color for gang activity.

When analyzing how much time SDPD spends on stops resulting in a field interview card, nearly six percent of all hours spent on officer-initiated stops are spent on stops that result in a field interview card.  

Within stops that result in a field interview card, SDPD disproportionately spends their time stopping people perceived as Latinx and Black. More than half of the hours spent on stops that result in a field interview card are spent stopping people perceived as Latinx or Black, even though Latinx and Black people make up only 35.7 percent of the total population in San Diego.

```{r item-chart, child= 'report_embeds/report_timespent_ficard_race_stop_item.Rmd'}
```

Black community members throughout San Diego are overwhelmingly impacted by this racially biased use of public resources. SDPD spends more time conducting field interview cards for Black people relative to their population in every police division. In Southeast San Diego, officers spent three times more hours on field interview cards of Black people compared to their relative share of the population in Southeast San Diego.  

```{r item-chart, child= 'report_embeds/report_timespent_div_ficard_race_person_group.Rmd'}
```

Analysis of hours spent by SDPD officers assigned to gang enforcement also show that most of their time is not being spent on activities that prevent crime. Offices assigned to gang enforcement spent nearly 4 in 10 hours on stops resulting only in a field interview card, a warning, or no action. Stops only resulting in a citation for an infraction took up an additional 8 out of 100 patrol hours. Comparatively, arrests with warrants only comprised 9 in 100 patrol hours.

```{r item-chart, child= 'report_embeds/report_timespent_ficard_gang_stop_item.Rmd'}
```

### B. This time translates to wasted public dollars.

`r lorem::ipsum(paragraphs = 1,sentences=NULL, avg_words_per_sentence=10)`


# RECOMMENDATIONS

1. Alternative investments that actually increase safety
2. Enact the Protect Act
3. Others...


# APPENDIX

```{r stacked-chart, child= 'report_embeds/report_stopsrates_age_race_gang_person_stacked.Rmd'}
```


```{r stacked-chart, child= 'report_embeds/report_stoprates_age_race_ficard_person_stacked.Rmd'}
```


# METHODOLOGY

[link to methodology on GitHub]