---
title: ""
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
---
<!-- link to Google Font -->
<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">
    
    
```{r, include = FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)

#### Load libraries ####
library(usethis)
library(tidyverse)
library(sf)
library(leaflet)
library(sp)
library(htmltools) # for html escape
library(magrittr)
library(RPostgreSQL)
library(htmlwidgets)
library(stringr)
  
source("W:\\RDA Team\\R\\credentials_source.R")
source("chart_functions.R")
  
pillars_conn <- connect_to_db("rjs_pillars")

```

```{r}
#### Map Data Prep ####
## load in race metadata
metadata_race_labels <-  dbGetQuery(pillars_conn, "SELECT * FROM metadata_race_labels") %>%
  select(race_base, race_label_short, race_label_long)
  
## Load in RJS Pillars data ##
df_long <- dbGetQuery(pillars_conn, "SELECT * FROM data.report_stoprates_div_ficard_race_person")
div <- st_read(pillars_conn, query =  "SELECT * FROM data.sangis_sdpd_divisions_2023") 
  
## change division name to sentence case
df_long$div_name <- str_to_title(df_long$div_name)
div$div_name <- str_to_title(div$div_name)
  
# transform for leaflet
div <- st_transform(div,4326)
  
# go from long to wide
df_wide <- df_long %>% 
  pivot_wider(names_from = race,
              values_from = total:rate,
              names_sep = "_")
  
# merge division sf by df: this is only for the shapefile
div_wide_sf <- div %>% 
  left_join(df_wide)
  
# the dot density needs the data to be long
div_long_sf <- div %>% 
  left_join(df_long)

## Calculate dot density counts
ficard_dots = as_dot_density(div_long_sf,
                             value = "count",
                             values_per_dot = 5, ## each dot represents 5 values
                             group = "race")

# Filter for race groups
df_long <- df_long %>% 
  group_by(race) %>% 
  summarize(count = n())

nh_black_dots <- ficard_dots %>% filter(race == "nh_black")
nh_white_dots <- ficard_dots %>% filter(race == "nh_white")
nh_aian_dots <- ficard_dots %>%  filter(race == "nh_aian")
latinx_dots <- ficard_dots %>% filter(race == "latinx")
sswana_dots <- ficard_dots %>% filter(race == "sswana")
nh_asian_dots <- ficard_dots %>% filter(race == "nh_asian")
nh_twoormor_dots <- ficard_dots %>% filter(race == "nh_twoormor")
nh_nhpi_dots <- ficard_dots %>% filter(race == "nh_nhpi")

invisible(dbDisconnect(pillars_conn))

```

```{r}
### Map components prep

## function to change the legend shape and color
colors_function <- function(colors, sizes, borders, shapes) {
    shapes <- gsub("circle", "100%", shapes)
    shapes <- gsub("square", "100%", shapes)
    paste0(colors, "; width:", sizes, "px; height:", 
           sizes, "px; border:1px solid ", borders, "; border-radius:", shapes)
}

## Create data_popup
# Stores ONLY data we plan to insert into the custom popup (i.e., not the full content like in rate_popup)
# Take note of the naming structure in the html tags: [data-][variable-name]
# It's important to only use hyphens (no underscores or other symbols to denote space between words)
# We'll use the names in the html tags when formatting popup behavior in JS (end of script)
data_popup <- paste0("<div class='pointdensity'> <div class='leaflet-data' data-division-name='", div_wide_sf$div_name,
               "' data-division-number='", div_wide_sf$div_num,
               "' data-count-nh-black='", div_wide_sf$count_nh_black,
               "' data-rate-nh-black='", round(div_wide_sf$rate_nh_black, digits = 1),
               "' data-count-latinx='", div_wide_sf$count_latinx,
               "' data-rate-latinx='", round(div_wide_sf$rate_latinx, digits = 1),
               "' data-count-nh-asian='", div_wide_sf$count_nh_asian,
               "' data-rate-nh-asian='", round(div_wide_sf$rate_nh_asian, digits = 1),
               "' data-count-nh-aian='", div_wide_sf$count_nh_aian,
               "' data-rate-nh-aian='", round(div_wide_sf$rate_nh_aian, digits = 1),
               "' data-count-nh-nhpi='", div_wide_sf$count_nh_nhpi,
               "' data-rate-nh-nhpi='", round(div_wide_sf$rate_nh_nhpi, digits = 1),
               "' data-count-sswana='", div_wide_sf$count_sswana,
               "' data-rate-sswana='", round(div_wide_sf$rate_sswana, digits = 1),
               "' data-count-nh-twoormor='", div_wide_sf$count_nh_twoormor,
               "' data-rate-nh-twoormor='", round(div_wide_sf$rate_nh_twoormor, digits = 1),
               "' data-count-nh-white='", div_wide_sf$count_nh_white,
               "' data-rate-nh-white='", round(div_wide_sf$rate_nh_white, digits = 1),
               "'></div></div>")


## Create pop-up content
# Stores full content (data and text) in an html string for the custom popup
# We are using <div> "class" names (e.g. popup-instruction, division-header) 
# to apply CSS styling (end of script, before JS)
# We are using <span> "class" names (e.g. data-emphasis, division-name) 
# to style with CSS AND insert the data stored in data_popup
# Note: the <span> "class" names are using the SAME naming as in the data_popup (only difference is no "data-" prefix) - this makes things clear and convenient for JS coding (end of script)

rate_popup <- "<div class='leaflet-popup-scrolled point-density-sidebar'><div class='popup-instruction'>Please click on a police division to see stop data below.</div><div class ='division-header'><span class='division-name'></span> Division (Division <span class='division-number'></span>)</div></div><br> SDPD Officers stopped and conducted field interviews on:<br><br>

<span class='nh-black-dot'></span> <span class='data-emphasis nh-black-color'><span class='count-nh-black nh-black-color'></span> people</span> perceived as <span class='data-emphasis nh-black-color'>Black</span>, or <span class='data-emphasis nh-black-color'><span class='rate-nh-black nh-black-color'></span>% of Black people stopped.</span><br><br> 

<span class='latinx-dot'></span> <span class='data-emphasis latinx-color'><span class='count-latinx latinx-color'></span> people</span> perceived as <span class='data-emphasis latinx-color'>Latinx</span>, or <span class='data-emphasis latinx-color'><span class='rate-latinx latinx-color'></span>% of Latinx people stopped.</span><br><br> 

<span class='nh-asian-dot'></span> <span class='nh-asian-color'><span class='count-nh-asian nh-asian-color'></span> people</span> perceived as <span class='data-emphasis nh-asian-color'>Asian</span>, or <span class='data-emphasis nh-asian-color'><span class='rate-nh-asian nh-asian-color'></span>% of Asian people stopped.</span><br><br> 

<span class='nh-aian-dot'></span> <span class='nh-aian-color'><span class='count-nh-aian nh-aian-color'></span> people</span> perceived as <span class='data-emphasis nh-aian-color'>AIAN</span>, or <span class='data-emphasis nh-aian-color'><span class='rate-nh-aian nh-aian-color'></span>% of AIAN people stopped.</span><br><br> 

<span class='nh-nhpi-dot'></span> <span class='nh-nhpi-color'><span class='count-nh-nhpi nh-nhpi-color'></span> people</span> perceived as <span class='nh-nhpi-color'>NHPI</span>, or <span class='nh-nhpi-color'><span class='rate-nh-nhpi'></span>% of NHPI people stopped.</span><br><br> 

<span class='sswana-dot'></span> <span class='sswana-color'><span class='count-sswana sswana-color'></span> people</span> perceived as <span class='sswana-color'>SWANA/SA</span>, or <span class='sswana-color'><span class='rate-sswana sswana-color'></span>% of SWANA/SA people stopped.</span><br><br> 

<span class='nh-twoormor-dot'></span> <span class='nh-twoormor-color'><span class='data-emphasis count-nh-twoormor nh-twoormor-color'></span> people perceived as <span class='data-emphasis'>Multiracial</span>, or <span class='rate-nh-twoormor data-emphasis nh-twoormor-color'></span>% of Multiracial people stopped.</span><br><br> 

<span class='nh-white-dot'></span> <span class='nh-white-color'><span class='data-emphasis count-nh-white nh-white-color'></span> people perceived as <span class='data-emphasis nh-white-color'>White</span>, or <span class='rate-nh-white data-emphasis nh-white-color'></span>% of White people stopped.</span><br><br> "

## Set CC Brand colors for race/ethnic layers
# Base Color 00
dark_tangerine <- c('#EB8D00') # AIAN
halaya_ube <- c('#601D42') # NHPI
american_yellow <- c('#EABB00') # SWANA/sa
halloween_orange <- c('#DF4007') # Asian
blue <- c('#0860BC') # Black
lime_green <- c('#0AB013') # Latinx
persian_indigo <- c('#211447') # Two or more races/Multiracial
bitter_lemon <- c('#BCD313') # White
  
# Create list of colors for race/ethnic layers
map_colors <- c(dark_tangerine, halaya_ube, american_yellow, halloween_orange,
                blue, lime_green, persian_indigo,  bitter_lemon)

# Create list of race/ethnic labels for map
map_race_labels <- c("AIAN",  "NHPI", "SWANA/SA", "Asian",  "Black", "Latinx", 
                     "Two or More Races", "White")

```
  
```{r}
### Build map
map <-  leaflet(width = "100%", height = "600px",
                options = leafletOptions(zoomControl = FALSE, minZoom=10, maxZoom=10)) %>%
  
  ## add provider tiles
  addProviderTiles("CartoDB.Positron") %>%
  
  ## add Division polygons
  # add polygon layer below (zIndex) markers - if the white fillColor is on top of the marker colors they will get washed out (dull)
  addPolygons(data = div_wide_sf, label=~htmlEscape(paste0(div_name, " Division ")),
              # polygon fill arguments
              fillColor="white", 
              #polygon line arguments
              color="black", weight=2, opacity = 1, smoothFactor = 0.5,
              # want to connect this to the data_popup (i.e. where we pull data values in)
              popup = ~paste0(data_popup),
              highlight = highlightOptions(color="darkgrey", weight = 2, opacity = 1,
                                           # Use these settings otherwise highlight will overlap/dull marker colors
                                           bringToFront=TRUE, sendToBack = TRUE)) %>%
  
  ### add NH White Stops
  addCircleMarkers(data = nh_white_dots, group = "White", label = NULL, 
                   weight = 2, opacity = 0.2, fillOpacity = 1, radius = 2, 
                   color="white",  fillColor=  bitter_lemon,  stroke = TRUE,
                   # keep this so that markers are layered on top of polygons 
                   # but do not impact our ability to click on the polygon
                   options = pathOptions(clickable = FALSE)) %>% 
  
  ### add nh two or more stops
  addCircleMarkers(data = nh_twoormor_dots, group = "Two or More Races", label = NULL, 
                   weight = 2, opacity = 0.2, fillOpacity = 1, radius = 2, 
                   color="white",  fillColor= persian_indigo, stroke = TRUE,
                   # keep this so that markers are layered on top of polygons 
                   # but do not impact our ability to click on the polygon
                   options = pathOptions(clickable = FALSE)) %>% 
  
  ### add latinx stops
  addCircleMarkers(data = latinx_dots, group = "Latinx", label = NULL, 
                   weight = 2, opacity = 0.2, fillOpacity = 1, radius = 2, 
                   color="white",  fillColor= lime_green,  stroke = TRUE,
                   # keep this so that markers are layered on top of polygons 
                   # but do not impact our ability to click on the polygon
                   options = pathOptions(clickable = FALSE)) %>% 
  
  ### add NH Black Stops
  addCircleMarkers(data = nh_black_dots, group = "Black", label = NULL, 
                   weight = 2, opacity = 0.2, fillOpacity = 1, radius = 2, 
                   color="white",  fillColor=blue, stroke = TRUE,
                   # keep this so that markers are layered on top of polygons 
                   # but do not impact our ability to click on the polygon
                   options = pathOptions(clickable = FALSE)) %>% 
  
  ### add nh asian stops
  addCircleMarkers(data = nh_asian_dots, group = "Asian", label = NULL, 
                   weight = 2, opacity = 0.2, fillOpacity = 1, radius = 2, 
                   color="white",  fillColor=  halloween_orange, stroke = TRUE,
                   # keep this so that markers are layered on top of polygons 
                   # but do not impact our ability to click on the polygon
                   options = pathOptions(clickable = FALSE)) %>% 
    
  ### add sswana stops
  addCircleMarkers(data = sswana_dots, group = "SWANA/SA", label = NULL, 
                   weight = 2, opacity = 0.2, fillOpacity = 1, radius = 2, 
                   color="white",  fillColor= american_yellow, stroke = TRUE,
                   # keep this so that markers are layered on top of polygons 
                   # but do not impact our ability to click on the polygon
                   options = pathOptions(clickable = FALSE)) %>%  
    
  ### add nh_nhpi stops
  addCircleMarkers(data = nh_nhpi_dots, group = "NHPI", label = NULL, 
                   weight = 2, opacity = 0.2, fillOpacity = 1, radius = 2, 
                   color="white",  fillColor= halaya_ube, stroke = TRUE,
                   # keep this so that markers are layered on top of polygons 
                   # but do not impact our ability to click on the polygon
                   options = pathOptions(clickable = FALSE)) %>% 
    
  ### add NH AIAN stops
  addCircleMarkers(data = nh_aian_dots,  group = "AIAN", label = NULL, 
                   weight = 2, opacity = 0.2, fillOpacity = 1, radius = 2, 
                   color="white",  fillColor= dark_tangerine, stroke = TRUE,
                   # keep this so that markers are layered on top of polygons 
                   # but do not impact our ability to click on the polygon
                   options = pathOptions(clickable = FALSE)) %>% 
  
  ### Layer control checkbox for race/ethnic groups
  addLayersControl(overlayGroups = map_race_labels, 
                   options = layersControlOptions(collapsed = FALSE)) %>%
    
  ### add Legend
  addLegend("bottomright", title= paste0("1 dot = 5 field interview card stops"), 
            colors = colors_function(colors = map_colors, sizes = "10", 
                                     shapes = "circle", borders = NULL),
            labels= map_race_labels, opacity = 0.7, labFormat = labelFormat(digits = 0)) %>% 
  
  ## add custom popup with full content
  addControl(html=rate_popup, position = "topleft") %>%
  
  ### set tiles and view
  addProviderTiles("CartoDB.Positron") %>%
  setView(-117.13395, 32.79923, zoom = 10) %>% #base and view 
    
  ### Deselect White from Layer Control checkbox
  hideGroup(c("White"))   # hide white group
  
```
  
<h3 class="leaflet-map-title">SDPD disproportionately conducts field interviews for people of color in the Southeastern and Central Division</h3>
<h4 class="leaflet-map-subtitle">People officers field interviewed by race and division</h4>
    
```{r}
  map
```
  <div align="left" style="line-height: 1em;"><span style=' font-size: 10px;'>Race/ethnicity: AIAN=American Indian or Alaska Native, NHPI=Native Hawaiian or Pacific Islander, SWANA/SA=Southwest Asian (Middle Eastern) or North African, or South Asian.</span> <span style="font-size: 10px;">Stops are randomly placed within each division and do not indicate the exact location of the stop.</span> <span style=' font-size: 10px;'>Catalyst California's calculations based on City of San Diego's Police Stop Data (2022); <u>[Catalyst California](https://www.catalystcalifornia.org)</u>, 2023. Analysis for all officer-initiated stops that resulted in a field interview card completed.</div>
    
<style type="text/css">

.division-header {
   margin-bottom: 8px; 
   font-weight: 700; 
   color: #000000;
   font-size: 14px; 
}

.leaflet-popup-scrolled {
  max-height: 580px;
  opacity: 1;
  border-bottom: 0;
  border-top: 0;
}

.leaflet-popup-scrolled .point-density-sidebar{
  max-width: 800px;
  max-height: 200px;
} 

.info {
  background: white;
  opacity: 1;
  border-radius: 0;
}

.leaflet-left {
  width: 33%;
}

.leaflet-left .leaflet-control {
  font-size: 12px;
  margin-left: 0;
  border-left: 1px solid #ddd;
  border-top: 1px solid #ddd;
  border-bottom: 1px solid #ddd;
}

.leaflet-right .leaflet-control {
  margin-right: 0;
}

.leaflet-top .leaflet-control {
  margin-top: 0;
}

.leaflet-control-container {
  color: #000000;
}

.hidden {
  display: none;
}

.popup-instruction {
  font-style: italic;
  font-weight: 600; 
  font-size: 1em; 
  line-height: 1.2em;
  color: #000000;
  border-bottom: 1px solid #000000;
  padding-top: 1.2em;
  padding-bottom: 1.2em;
  margin-bottom: 1.2em;
}

.legend {
  color: #000000;
  border-top: 1px solid #000000;
  padding-top: 1.2em;
}

.legend-note {
  font-size: 1em;
  font-weight: 700;
}

.nh-black-dot {
  height: 6px;
  width: 6px;
  background-color: #0860BC;
  border-radius: 50%;
  display: inline-block;
}

.nh-black-color {
  color: #0860BC;
}

.latinx-dot {
  height: 6px;
  width: 6px;
  background-color: #0AB013;
  border-radius: 50%;
  display: inline-block;
}

.latinx-color {
  color: #0AB013;
}

.nh-asian-dot {
  height: 6px;
  width: 6px;
  background-color: #DF4007;
  border-radius: 50%;
  display: inline-block;
}

.nh-asian-color {
  color: #DF4007;
}

.nh-aian-dot {
  height: 6px;
  width: 6px;
  background-color: #EB8D00;
  border-radius: 50%;
  display: inline-block;
}

.nh-aian-color {
  color: #EB8D00;
}

.nh-nhpi-dot {
  height: 6px;
  width: 6px;
  background-color: #601D42;
  border-radius: 50%;
  display: inline-block;
}

.nh-nhpi-color {
  color: #601D42;
}

.sswana-dot {
  height: 6px;
  width: 6px;
  background-color: #EABB00;
  border-radius: 50%;
  display: inline-block;
}

.sswana-color {
  color: #EABB00;
}

.nh-twoormor-dot {
  height: 6px;
  width: 6px;
  background-color: #211447;
  border-radius: 50%;
  display: inline-block;
}

.nh-twoormor-color {
  color: #211447;
}

.nh-white-dot {
  height: 6px;
  width: 6px;
  background-color: #BCD313;
  border-radius: 50%;
  display: inline-block;
}

.nh-white-color {
  color: #BCD313;
}

.data-emphasis {
  font-size: 14px; 
  font-weight: 600;
}

</style>

```{js}
var mapsPlaceholder = [];
L.Map.addInitHook(function () {
  mapsPlaceholder.push(this);
  mapsPlaceholder.forEach(map => {
    // pre-populate map pop-up (when map loads/before user clicks anywhere)
    map.on('load', (e) => {
    setTimeout(() => {
    document.querySelector('.division-name').innerHTML = 'Southeastern'
    document.querySelector('.division-number').innerHTML = '4'
    document.querySelector('.count-nh-black').innerHTML = '549'
    document.querySelector('.rate-nh-black').innerHTML = '19.4'
    document.querySelector('.count-latinx').innerHTML = '405'
    document.querySelector('.rate-latinx').innerHTML = '11.0'
    document.querySelector('.count-nh-asian').innerHTML = '42'
    document.querySelector('.rate-nh-asian').innerHTML = '10.2'
    document.querySelector('.count-nh-aian').innerHTML = '4'
    document.querySelector('.rate-nh-aian').innerHTML = '25'
    document.querySelector('.count-nh-nhpi').innerHTML = '7'
    document.querySelector('.rate-nh-nhpi').innerHTML = '13.2'
    document.querySelector('.count-sswana').innerHTML = '3'
    document.querySelector('.rate-sswana').innerHTML = '3.0'
    document.querySelector('.count-nh-twoormor').innerHTML = '8'
    document.querySelector('.rate-nh-twoormor').innerHTML = '21.6'
    document.querySelector('.count-nh-white').innerHTML = '82'
    document.querySelector('.rate-nh-white').innerHTML = '9.6'
    }, 250)
  })
    // hide default leaflet popup
    map.on('popupopen', (e) => {
    let popups = document.querySelectorAll('.leaflet-data');
    // console.log(popups);

    document.querySelectorAll('.pointdensity').forEach(pointdensity => {
    pointdensity.parentElement.parentElement.parentElement.classList.add('hidden')
  })
  
  // populate custom pop-up when user clicks map
  //console.log('click')
  let popup = popups[popups.length-1]
  document.querySelector('.division-name').innerHTML = popup.getAttribute('data-division-name')
  document.querySelector('.division-number').innerHTML = popup.getAttribute('data-division-number')
  document.querySelector('.count-nh-black').innerHTML = popup.getAttribute('data-count-nh-black')
  document.querySelector('.rate-nh-black').innerHTML = popup.getAttribute('data-rate-nh-black')
  document.querySelector('.count-latinx').innerHTML = popup.getAttribute('data-count-latinx')
  document.querySelector('.rate-latinx').innerHTML = popup.getAttribute('data-rate-latinx')
  document.querySelector('.count-nh-asian').innerHTML = popup.getAttribute('data-count-nh-asian')
  document.querySelector('.rate-nh-asian').innerHTML = popup.getAttribute('data-rate-nh-asian')
  document.querySelector('.count-nh-aian').innerHTML = popup.getAttribute('data-count-nh-aian')
  document.querySelector('.rate-nh-aian').innerHTML = popup.getAttribute('data-rate-nh-aian')
  document.querySelector('.count-nh-nhpi').innerHTML = popup.getAttribute('data-count-nh-nhpi')
  document.querySelector('.rate-nh-nhpi').innerHTML = popup.getAttribute('data-rate-nh-nhpi')
  document.querySelector('.count-sswana').innerHTML = popup.getAttribute('data-count-sswana')
  document.querySelector('.rate-sswana').innerHTML = popup.getAttribute('data-rate-sswana')
  document.querySelector('.count-nh-twoormor').innerHTML = popup.getAttribute('data-count-nh-twoormor')
  document.querySelector('.rate-nh-twoormor').innerHTML = popup.getAttribute('data-rate-nh-twoormor')
  document.querySelector('.count-nh-white').innerHTML = popup.getAttribute('data-count-nh-white')
  document.querySelector('.rate-nh-white').innerHTML = popup.getAttribute('data-rate-nh-white')
})})});

```  