---
title: ""
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
---
<!-- link to Google Font -->
<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">
    
    
```{r, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)

#### Load libraries ####
library(usethis)
library(tidyverse)
library(sf)
library(leaflet)
library(sp)
library(htmltools) # for html escape
library(magrittr)
library(RPostgreSQL)
library(htmlwidgets)
library(stringr)
  
source("W:\\RDA Team\\R\\credentials_source.R")
source("chart_functions.R")
  
pillars_conn <- connect_to_db("rjs_pillars")
```

```{r}
#### Map: Data Prep ###
## load in race metadata
metadata_race_labels <-  dbGetQuery(pillars_conn, "SELECT * FROM metadata_race_labels") %>% 
  select(race_base, race_label_short, race_label_long)
  
## Load in RJS Pillars data ##
df_long <- dbGetQuery(pillars_conn, "SELECT * FROM data.report_hitrates_div_race_ficard_person") %>% 
  filter(race == "total")

div <- st_read(pillars_conn, query =  "SELECT * FROM data.sangis_sdpd_divisions_2023") 
  
## change division name to sentence case
df_long$div_name <- str_to_title(df_long$div_name)
div$div_name <- str_to_title(div$div_name)
  
## add contraband count/contraband prc rate
df_long <- df_long %>% 
  mutate(contraband_count = (stop_count-no_contraband_count), 
         contraband_prc_rate = (100-no_contraband_prc_rate))
  
# transform for leaflet
div <- st_transform(div, 4326)
  
# go from long to wide
df_wide <- df_long %>% 
  pivot_wider(names_from = race,
              values_from = stop_count:no_contraband_prc_rate,
              names_sep = "_")
  
# merge division sf by df: this is only for the shapefile
div_wide_sf <- div %>% 
  left_join(df_wide)
  
  
# the dot density needs the data to be long
div_long_sf <- div %>% 
  left_join(df_long)

### Calculate dot density count ###
no_contraband_dots = as_dot_density(div_long_sf,
                                    value = "no_contraband_count",
                                    values_per_dot = 5, ## each dot represents
                                    group = "race")
    
contraband_dots = as_dot_density(div_long_sf,
                                 value = "contraband_count",
                                 values_per_dot = 5, ## each dot represents
                                 group = "race")
```

```{r}
### Map components prep

## Set CC Brand Colors for contraband layers
# Base Color 00
meteorite <- c('#3A207D') # no contraband
papaya <- c('#F25922') # contraband

## function to change the legend shape and color
colors_function <- function(colors, sizes, borders, shapes) {
    shapes <- gsub("circle", "100%", shapes)
    shapes <- gsub("square", "100%", shapes)
    paste0(colors, "; width:", sizes, "px; height:", sizes, "px; border:1px solid ", borders, "; border-radius:", shapes)}

## Create data content for popup
# This stores the values we plan to insert into the custom popup (i.e., not the full content of the custom popup like in rate_popup)
# Take note of the naming structure in the html tags: [data-][variable-name]
# It's important to only use hyphens (no underscores or other symbols to denote space between words)
# We'll use the names in the html tags when formatting popup behavior in JS (end of script)
data_popup <- paste0("<div class='pointdensity'> <div class='leaflet-data' data-division-name='", div_wide_sf$div_name,
               "' data-division-number='", div_wide_sf$div_num,
               "' data-stop-count-total='", div_wide_sf$stop_count_total,
               "' data-yes-contraband-count='", div_wide_sf$contraband_count,
               "' data-yes-contraband-prc-rate='", round(div_wide_sf$contraband_prc_rate, digits = 1),
               "' data-no-contraband-count='", div_wide_sf$no_contraband_count_total,
               "' data-no-contraband-prc-rate='", round(div_wide_sf$no_contraband_prc_rate_total, digits = 1),
               "' data-division-name-default='", div_wide_sf$div_name[1],
               "' data-division-number-default='", div_wide_sf$div_num[1],
               "' data-stop-count-total-default='", div_wide_sf$stop_count_total[1],
               "' data-yes-contraband-count-default='", div_wide_sf$contraband_count[1],
               "' data-yes-contraband-prc-rate-default='", round(div_wide_sf$contraband_prc_rate[1], digits = 1),
               "' data-no-contraband-count-default='", div_wide_sf$no_contraband_count_total[1],
               "' data-no-contraband-prc-rate-default='", round(div_wide_sf$no_contraband_prc_rate_total[1], digits = 1),
               "'></div></div>")

## Create full content (data and text) in an html string for the custom popup
# Important to note "class" names (e.g., leaflet-popup-scrolled, division-header, stop-count-total, etc.) 
# <div> "class" names are used as references with CSS styling (end of script, before JS)
# <span> "class" names help us with CSS AND inserting the data we have stored in data_popup
# Note: the <span> "class" names are using the SAME naming as in the data_popup (only difference is no "data-" prefix) - this is a convenience for the JS (end of script)
rate_popup <- "<div class='leaflet-popup-scrolled point-density-sidebar'><div class='popup-instruction'>Please click on a police division to see stop data below.</div><div class ='division-header'><span class='division-name'></span> Division (Division <span class='division-number'></span>)</div></div><br>Out of <span class='data-emphasis'><span class='stop-count-total'></span> people</span> searched during stops where a field interview card was completed:<br><br>SDPD found no contraband or evidence on <span class='data-emphasis no-contraband-color'><span class='no-contraband-count no-contraband-color'></span> people</span>, meaning SDPD  subjected <span class='data-emphasis no-contraband-color'><span class='no-contraband-prc-rate no-contraband-color'></span>%</span> of people to unnecessary searches.<br><br>SDPD found contraband/evidence on <span class='data-emphasis yes-contraband-color'><span class='yes-contraband-count yes-contraband-color'></span> people</span>, a hit rate of <span class='data-emphasis yes-contraband-color'><span class='yes-contraband-prc-rate yes-contraband-color'></span>%</span>.
<br><br><div class='legend'><div class='legend-note'>1 dot = 5 field interview card stops and a search</div><br><span class='no-contraband-dot'></span> No Contraband Found <br><span class='yes-contraband-dot'></span> Yes Contraband Found</div></div><br>"
```

```{r}

### Build map
map <- leaflet(width = "100%", height = "600px",
               options = leafletOptions(zoomControl = FALSE, minZoom=10, maxZoom=10)) %>%
  ## add provider tiles
  addProviderTiles("CartoDB.Positron") %>%
  
  ## add Division polygons
  # add polygon layer below (zIndex) markers - if the white fillColor is on top of the marker colors they will get washed out (dull)
  addPolygons(data = div_wide_sf, label=~htmlEscape(paste0(div_name, " Division ")),
              # polygon fill arguments
              fillColor="white", 
              #polygon line arguments
              color="black", weight=2, opacity = 1, smoothFactor = 0.5,
              # want to connect this to the data_popup (i.e. where we pull data values in)
              popup = ~paste0(data_popup),
              highlight = highlightOptions(color="darkgrey", weight = 2, opacity = 1,
                                           # Use these settings otherwise highlight will overlap/change marker colors
                                           bringToFront=TRUE, sendToBack = TRUE)) %>%
  
  ## add no contraband stop markers
  addCircleMarkers(data = no_contraband_dots, group = "No Contraband Found",
                   label = NULL, weight = 3,  opacity = 0, fillOpacity = 1, 
                   radius = 2, fillColor= papaya, stroke = TRUE,
                   # keep this so that markers are layered on top of polygons 
                   # but do not impact our ability to click on the polygon
                   options = pathOptions(clickable = FALSE)) %>% 
    
  ## add contraband stop markers
  addCircleMarkers(data = contraband_dots, group = "Yes Contraband Found",
                   label = NULL, weight = 3, opacity = 0, fillOpacity = 1,
                   radius = 2, fillColor=meteorite,  stroke = TRUE,
                   # keep this so that markers are layered on top of polygons 
                   # but do not impact our ability to click on the polygon
                   options = pathOptions(clickable = FALSE)) %>%

  ## Add selection control for contraband/no contraband layers
  addLayersControl(overlayGroups = c("No Contraband Found", "Yes Contraband Found"),
                   options = layersControlOptions(collapsed = FALSE)) %>%
    
  ## add custom popup with full content
  addControl(html=rate_popup, position = "topleft") %>%

  ## Set view
  setView(-117.13395, 32.79923, zoom = 10)
  
```
  
<h3 class="leaflet-map-title">SDPD conducts the highest number of unnecessary searches during field interview card stops in the Southeastern Division</h3>
<h4 class="leaflet-map-subtitle">People searched during field interviews by division and search result</h4>
    
    
```{r}
map
```
<div align="left" style="line-height: 1em;"><span style="font-size: 10px;">Searches are randomly placed within each division and do not indicate the exact location of the search.</span> <span style=' font-size: 10px;'>Catalyst California's calculations based on City of San Diego's Police Stop Data (2022); <u>[Catalyst California](https://www.catalystcalifornia.org)</u>, 2023. Analysis for all officer-initiated stops that resulted in a field interview card completed and a search.</span> </div>


<style type="text/css">

.division-header {
   margin-bottom: 8px; 
   font-weight: 700; 
   color: #000000;
   font-size: 14px; 
}

.no-contraband-color {
  color: #F25922;
}

.yes-contraband-color {
  color: #3A207D;
}

.leaflet-popup-scrolled {
  max-height: 580px;
  opacity: 1;
  border-bottom: 0;
  border-top: 0;
}

.leaflet-popup-scrolled .point-density-sidebar{
  max-width: 800px;
  max-height: 200px;
} 

.info {
  background: white;
  opacity: 1;
  border-radius: 0;
}

.leaflet-left {
  width: 33%;
}

.leaflet-left .leaflet-control {
  font-size: 12px;
  margin-left: 0;
  border-left: 1px solid #ddd;
  border-top: 1px solid #ddd;
  border-bottom: 1px solid #ddd;
}

.leaflet-right .leaflet-control {
  margin-right: 0;
}

.leaflet-top .leaflet-control {
  margin-top: 0;
}

.leaflet-control-container {
  color: #000000;
}

.hidden {
  display: none;
}

.popup-instruction {
  font-style: italic;
  font-weight: 600; 
  font-size: 1em; 
  line-height: 1.2em;
  color: #000000;
  border-bottom: 1px solid #000000;
  padding-top: 1.2em;
  padding-bottom: 1.2em;
  margin-bottom: 1.2em;
}

.legend {
  color: #000000;
  border-top: 1px solid #000000;
  padding-top: 1.2em;
}

.legend-note {
  font-size: 1em;
  font-weight: 700;
}

.yes-contraband-dot {
  height: 1em;
  width: 1em;
  background-color: #3A207D;
  border-radius: 50%;
  display: inline-block;
}


.no-contraband-dot {
  height: 1em;
  width: 1em;
  background-color: #F25922;
  border-radius: 50%;
  display: inline-block;
}

.data-emphasis {
  font-size: 14px; 
  font-weight: 700;
}

</style>

```{js}

 
var mapsPlaceholder = [];
L.Map.addInitHook(function () {
  mapsPlaceholder.push(this);
  mapsPlaceholder.forEach(map => {
    map.on('load', (e) => {
    setTimeout(() => {
    document.querySelector('.division-name').innerHTML = 'Northern'
    document.querySelector('.division-number').innerHTML = '1'
    document.querySelector('.stop-count-total').innerHTML = '312'
    document.querySelector('.yes-contraband-count').innerHTML = '66'
    document.querySelector('.yes-contraband-prc-rate').innerHTML = '21.2'
    document.querySelector('.no-contraband-count').innerHTML = '246'
    document.querySelector('.no-contraband-prc-rate').innerHTML = '78.8'
    }, 250)
  })
    map.on('popupopen', (e) => {
    let popups = document.querySelectorAll('.leaflet-data');
    // console.log(popups);

    document.querySelectorAll('.pointdensity').forEach(pointdensity => {
    pointdensity.parentElement.parentElement.parentElement.classList.add('hidden')
  })
  
  //console.log('click')
  let popup = popups[popups.length-1]
  document.querySelector('.division-name').innerHTML = popup.getAttribute('data-division-name')
  document.querySelector('.division-number').innerHTML = popup.getAttribute('data-division-number')
  document.querySelector('.stop-count-total').innerHTML = popup.getAttribute('data-stop-count-total')
  document.querySelector('.yes-contraband-count').innerHTML = popup.getAttribute('data-yes-contraband-count')
  document.querySelector('.yes-contraband-prc-rate').innerHTML = popup.getAttribute('data-yes-contraband-prc-rate')
  document.querySelector('.no-contraband-count').innerHTML = popup.getAttribute('data-no-contraband-count')
  document.querySelector('.no-contraband-prc-rate').innerHTML = popup.getAttribute('data-no-contraband-prc-rate')

})})});

```
