---
title: ""
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
---
<!-- link to Google Font -->
<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">
    
    
```{r, include=FALSE}
#### Map: Data Prep ###
#### Load libraries ####
library(usethis)
library(tidyverse)
library(sf)
library(leaflet)
library(sp)
library(htmltools) # for html escape
library(magrittr)
library(RPostgreSQL)
library(htmlwidgets)
library(stringr)
  
source("W:\\RDA Team\\R\\credentials_source.R")
source("chart_functions.R")
  
pillars_conn <- connect_to_db("rjs_pillars")
  
## load in race metadata
metadata_race_labels <-  dbGetQuery(pillars_conn, "SELECT * FROM metadata_race_labels") %>% 
  select(race_base, race_label_short, race_label_long)
  
## Load in RJS Pillars data ##
df_long <- dbGetQuery(pillars_conn, "SELECT * FROM data.report_hitrates_div_race_ficard_person") %>% 
  filter(race == "total")

div <- st_read(pillars_conn, query =  "SELECT * FROM data.sangis_sdpd_divisions_2023") 
  
## change division name to sentence case
df_long$div_name <- str_to_title(df_long$div_name)
div$div_name <- str_to_title(div$div_name)
  
## add contraband count/contraband prc rate
df_long <- df_long %>% 
  mutate(contraband_count = (stop_count-no_contraband_count), 
         contraband_prc_rate = (100-no_contraband_prc_rate))
  
# transform for leaflet
div <- st_transform(div, 4326)
  
# go from long to wide
df_wide <- df_long %>% 
  pivot_wider(names_from = race,
              values_from = stop_count:no_contraband_prc_rate,
              names_sep = "_")
  
# merge division sf by df: this is only for the shapefile
div_wide_sf <- div %>% 
  left_join(df_wide)
  
  
# the dot density needs the data to be long
div_long_sf <- div %>% 
  left_join(df_long)
```
  

```{r, include=FALSE}
### Calculate dot density count ###
no_contraband_dots = as_dot_density(div_long_sf,
                                    value = "no_contraband_count",
                                    values_per_dot = 5, ## each dot represents
                                    group = "race")
    
contraband_dots = as_dot_density(div_long_sf,
                                 value = "contraband_count",
                                 values_per_dot = 5, ## each dot represents
                                 group = "race")

```
  
  
```{r, include=FALSE}
### Map components prep

## Set CC Brand Colors for contraband layers
# Base Color 00
meteorite <- c('#3A207D') # no contraband
peridot <- c('#F25922') # contraband

## function to change the legend shape and color
colors_function <- function(colors, sizes, borders, shapes) {
    shapes <- gsub("circle", "100%", shapes)
    shapes <- gsub("square", "100%", shapes)
    paste0(colors, "; width:", sizes, "px; height:", sizes, "px; border:1px solid ", borders, "; border-radius:", shapes)}

## Create content for popup
# This stores the values we plan to insert into the final popup (i.e., not the full content of the final popup like rate_popup)
# Take note of the naming structure in the html tags: [data-][variable-name]
# It's important to only use hyphens (no underscores or other symbols to denote space between words)
# We'll use the names in the html tags when formatting popup behavior in JS (end of script)
data_popup <- paste0("<div class='pointdensity'> <div class='leaflet-data' data-division-name='", div_wide_sf$div_name,
               "' data-division-number='", div_wide_sf$div_num,
               "' data-stop-count-total='", div_wide_sf$stop_count_total,
               "' data-contraband-count='", div_wide_sf$contraband_count,
               "' data-contraband-prc-rate='", round(div_wide_sf$contraband_prc_rate, digits = 1),
               "' data-no-contraband-count-total='", div_wide_sf$no_contraband_count_total,
               "' data-no-contrband-prc-rate-total='", round(div_wide_sf$no_contraband_prc_rate_total, digits = 1),
               "' ></div></div>")

# rate_popup <- paste0("<div class='leaflet-popup-scrolled' style='max-width:800px;max-height:200px'>",
#                      "In the <span style='font-size: 14px; font-weight: bold;'>",
#                      div_wide_sf$div_name, " Division",  " (Division ", div_wide_sf$div_num,")",
#                      "</span>", "<br> <br> Out of ", "<span style='font-size: 14px; font-weight: bold;'>", div_wide_sf$stop_count_total,
#                      "</span> people searched during stops where a field interview card was completed, SDPD found contraband/evidence on ",
#                      "<span style='font-size: 14px; font-weight: bold;'>", div_wide_sf$contraband_count, 
#                      "</span>", " people, a hit rate of ", "<span style='font-size: 14px; font-weight: bold;'>",
#                      round(div_wide_sf$contraband_prc_rate, digits = 1), "</span>",  "%.", 
#                      " SDPD found no contraband or evidence on ", "<span style='font-size: 14px; font-weight: bold;'>",
#                      div_wide_sf$no_contraband_count_total, "</span>", " people, meaning SDPD  subjected ", 
#                      "<span style='font-size: 14px; font-weight: bold;'>", round(div_wide_sf$no_contraband_prc_rate_total, digits = 1),
#                      "% of people to unnecessary searches.")
#   
```
  
  
```{r, include=FALSE}
### Build map
map <- leaflet(width = "100%", height = "600px") %>%
    
  ## add no contraband stop markers
  addCircleMarkers(data = no_contraband_dots, group = "No Contraband Found",
                   label = NULL, weight = 3,  opacity = 0, fillOpacity = 1, 
                   radius = 2, color="white",  fillColor= meteorite, stroke = TRUE) %>% 
    
  ## add contraband stop markers
  addCircleMarkers(data = contraband_dots, group = "Yes Contraband Found",
                   label = NULL, weight = 3, opacity = 0, fillOpacity = 1,
                   radius = 2, color="white",  fillColor=  peridot,  stroke = TRUE) %>% 
    
  ## add Division polygons
  addPolygons(data = div_wide_sf, label=~htmlEscape(paste0(div_name, " Division ")),
              color = "black", fillColor = "white", opacity = 1,  popup = ~paste0(rate_popup),
              weight = 2, smoothFactor = 0.5, highlight = highlightOptions(color = "white", 
                                                                           weight = 3, bringToFront = TRUE, sendToBack = FALSE)) %>%

  ## Add selection control for contraband/no contraband layers
  addLayersControl(overlayGroups = c("No Contraband Found", "Yes Contraband Found"),
                   options = layersControlOptions(collapsed = FALSE)) %>%
    
  ## add Legend
  addLegend(position="bottomleft", title=paste0("1 dot = 5 field interview card stops and a search"),
            colors = colors_function(colors = c(meteorite, peridot), sizes = "10", shapes = "circle", borders = NULL),
            labels= c("No Contraband Found",  "Yes Contraband Found"),
            opacity = 0.7, labFormat = labelFormat(digits = 0)) %>% 
  
  addProviderTiles("CartoDB.Positron") %>%
  setView(-117.13395, 32.79923, zoom = 10)
  
```
  
<h3 class="leaflet-map-title">SDPD conducts the highest number of unnecessary searches during field interview card stops in the Southeastern Division</h3>
<h4 class="leaflet-map-subtitle">People searched during field interviews by division and search result</h4>
    
    
```{r, echo=FALSE}
  map
```
<div align="left" style="line-height: 1em;"><span style="font-size: 10px;">Searches are randomly placed within each division and do not indicate the exact location of the search.</span> <span style=' font-size: 10px;'>Catalyst California's calculations based on City of San Diego's Police Stop Data (2022); <u>[Catalyst California](https://www.catalystcalifornia.org)</u>, 2023. Analysis for all officer-initiated stops that resulted in a field interview card completed and a search.</span> </div>
    