---
title: ""
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
---
<!-- link to Google Font -->
<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">
    
    
```{r, include = FALSE}

#### Load libraries ####
library(usethis)
library(tidyverse)
library(sf)
library(leaflet)
library(sp)
library(htmltools) # for html escape
library(magrittr)
library(RPostgreSQL)
library(htmlwidgets)
library(stringr)
  
source("W:\\RDA Team\\R\\credentials_source.R")
source("chart_functions.R")
  
pillars_conn <- connect_to_db("rjs_pillars")

#### Map Data Prep ####
## load in race metadata
metadata_race_labels <-  dbGetQuery(pillars_conn, "SELECT * FROM metadata_race_labels") %>%
  select(race_base, race_label_short, race_label_long)
  
## Load in RJS Pillars data ##
df_long <- dbGetQuery(pillars_conn, "SELECT * FROM data.report_stoprates_div_ficard_race_person")
div <- st_read(pillars_conn, query =  "SELECT * FROM data.sangis_sdpd_divisions_2023") 
  
## change division name to sentence case
df_long$div_name <- str_to_title(df_long$div_name)
div$div_name <- str_to_title(div$div_name)
  
# transform for leaflet
div <- st_transform(div,4326)
  
# go from long to wide
df_wide <- df_long %>% 
  pivot_wider(names_from = race,
              values_from = total:rate,
              names_sep = "_")
  
# merge division sf by df: this is only for the shapefile
div_wide_sf <- div %>% 
  left_join(df_wide)
  
# the dot density needs the data to be long
div_long_sf <- div %>% 
  left_join(df_long)

## Calculate dot density counts
ficard_dots = as_dot_density(div_long_sf,
                             value = "count",
                             values_per_dot = 5, ## each dot represents 5 values
                             group = "race")

# Filter for race groups
df_long %>% 
  as.data.frame() %>% 
  group_by(race) %>% 
  summarize(count = n())

nh_black_dots <- ficard_dots %>% filter(race == "nh_black")
nh_white_dots <- ficard_dots %>% filter(race == "nh_white")
nh_aian_dots <- ficard_dots %>%  filter(race == "nh_aian")
latinx_dots <- ficard_dots %>% filter(race == "latinx")
sswana_dots <- ficard_dots %>% filter(race == "sswana")
nh_asian_dots <- ficard_dots %>% filter(race == "nh_asian")
nh_twoormor_dots <- ficard_dots %>% filter(race == "nh_twoormor")
nh_nhpi_dots <- ficard_dots %>% filter(race == "nh_nhpi")

dbDisconnect(pillars_conn)
```

```{r, include=FALSE}
### Map components prep

## function to change the legend shape and color
colors_function <- function(colors, sizes, borders, shapes) {
    shapes <- gsub("circle", "100%", shapes)
    shapes <- gsub("square", "100%", shapes)
    paste0(colors, "; width:", sizes, "px; height:", 
           sizes, "px; border:1px solid ", borders, "; border-radius:", shapes)
}

## Create data_popup
# Stores ONLY data we plan to insert into the custom popup (i.e., not the full content like in rate_popup)
# Take note of the naming structure in the html tags: [data-][variable-name]
# It's important to only use hyphens (no underscores or other symbols to denote space between words)
# We'll use the names in the html tags when formatting popup behavior in JS (end of script)
data_popup <- paste0("<div class='pointdensity'> <div class='leaflet-data' data-division-name='", div_wide_sf$div_name,
               "' data-division-number='", div_wide_sf$div_num,
               "' data-count-nh-black='", div_wide_sf$count_nh_black,
               "' data-rate-nh-black='", round(div_wide_sf$rate_nh_black, digits = 1),
               "' data-count-latinx='", div_wide_sf$count_nh_black,
               "' data-rate-latinx='", round(div_wide_sf$rate_nh_black, digits = 1),
               "' data-count-nh-asian='", div_wide_sf$count_nh_black,
               "' data-rate-nh-asian='", round(div_wide_sf$rate_nh_black, digits = 1),
               "' data-count-nh-aian='", div_wide_sf$count_nh_black,
               "' data-rate-nh-aian='", round(div_wide_sf$rate_nh_black, digits = 1),
               "' data-count-nh-nhpi='", div_wide_sf$count_nh_black,
               "' data-rate-nh-nhpi='", round(div_wide_sf$rate_nh_black, digits = 1),
               "' data-count-sswana='", div_wide_sf$count_nh_black,
               "' data-rate-sswana='", round(div_wide_sf$rate_nh_black, digits = 1),
               "' data-count-nh-twoormor='", div_wide_sf$count_nh_black,
               "' data-rate-nh-twoormor='", round(div_wide_sf$rate_nh_black, digits = 1),
               "' data-count-nh-white='", div_wide_sf$count_nh_black,
               "' data-rate-nh-white='", round(div_wide_sf$rate_nh_black, digits = 1),
               "'></div></div>")


## Create pop-up content
# Stores full content (data and text) in an html string for the custom popup
# We are using <div> "class" names (e.g. popup-instruction, division-header) 
# to apply CSS styling (end of script, before JS)
# We are using <span> "class" names (e.g. data-emphasis, division-name) 
# to style with CSS AND insert the data stored in data_popup
# Note: the <span> "class" names are using the SAME naming as in the data_popup (only difference is no "data-" prefix) - this makes things clear and convenient for JS coding (end of script)

rate_popup <- "<div class='leaflet-popup-scrolled point-density-sidebar'><div class='popup-instruction'>Please click on a police division to see stop data below.</div><div class ='division-header'><span class='division-name'></span> Division (Division <span class='division-number'></span>)</div></div><br><br>

Officers stopped <span class='count-nh-black'></span> people perceived as <span class='data-emphasis'>Black</span> for stops that resulted in a field interview card completed, or<span class='rate-nh-black'></span>% of Black people stopped.<br><br> 

Officers stopped <span class='count-latinx'></span> people perceived as <span class='data-emphasis'>Latinx</span> for stops that resulted in a field interview card completed, or<span class='rate-latinx'></span>% of Latinx people stopped.<br><br> 

Officers stopped <span class='count-nh-asian'></span> people perceived as <span class='data-emphasis'>Asian</span> for stops that resulted in a field interview card completed, or<span class='rate-nh-asian'></span>% of Asian people stopped.<br><br> 

Officers stopped <span class='count-nh-aian'></span> people perceived as <span class='data-emphasis'>AIAN</span> for stops that resulted in a field interview card completed, or<span class='rate-nh-aian'></span>% of AIAN people stopped.<br><br> 

Officers stopped <span class='count-nh-nhpi'></span> people perceived as <span class='data-emphasis'>NHPI</span> for stops that resulted in a field interview card completed, or<span class='rate-nh-nhpi'></span>% of NHPI people stopped.<br><br> 

Officers stopped <span class='count-sswana'></span> people perceived as <span class='data-emphasis'>SWANA/SA</span> for stops that resulted in a field interview card completed, or<span class='rate-sswana'></span>% of SWANA/SA people stopped.<br><br> 

Officers stopped <span class='count-nh-twoormor'></span> people perceived as <span class='data-emphasis'>Multiracial</span> for stops that resulted in a field interview card completed, or<span class='rate-nh-twoormor'></span>% of Multiracial people stopped.<br><br> 

Officers stopped <span class='count-nh-white'></span> people perceived as <span class='data-emphasis'>White</span> for stops that resulted in a field interview card completed, or<span class='rate-nh-white'></span>% of White people stopped.<br><br> "

## Set CC Brand colors for race/ethnic layers
# Base Color 00
lime_green <- c('#0AB013')
persian_indigo <- c('#211447')
halloween_orange <- c('#DF4007')
bitter_lemon <- c('#BCD313')
blue <- c('#0860BC')
dark_tangerine <- c('#EB8D00')
american_yellow <- c('#EABB00')
halaya_ube <- c('#601D42')
  
# Create list of colors for race/ethnic layers
map_colors <- c(dark_tangerine, halaya_ube, american_yellow, halloween_orange,
                blue, lime_green, persian_indigo,  bitter_lemon)

# Create list of race/ethnic labels for map
map_race_labels <- c("AIAN",  "NHPI", "SWANA/SA", "Asian",  "Black", "Latinx", 
                     "Two or More Races", "White")

```
  
```{r, include=FALSE}
### Build map
map <-  leaflet(width = "100%", height = "600px") %>%
  
  ### add NH White Stops
  addCircleMarkers(data = nh_white_dots, group = "White", label = NULL, 
                   weight = 2, opacity = 0.2, fillOpacity = 1, radius = 2, 
                   color="white",  fillColor=  bitter_lemon,  stroke = TRUE) %>%
  
  ### add nh two or more stops
  addCircleMarkers(data = nh_twoormor_dots, group = "Two or More Races", label = NULL, 
                   weight = 2, opacity = 0.2, fillOpacity = 1, radius = 2, 
                   color="white",  fillColor= persian_indigo, stroke = TRUE) %>% 
  
  ### add latinx stops
  addCircleMarkers(data = latinx_dots, group = "Latinx", label = NULL, 
                   weight = 2, opacity = 0.2, fillOpacity = 1, radius = 2, 
                   color="white",  fillColor= lime_green,  stroke = TRUE) %>%
  
  ### add NH Black Stops
  addCircleMarkers(data = nh_black_dots, group = "Black", label = NULL, 
                   weight = 2, opacity = 0.2, fillOpacity = 1, radius = 2, 
                   color="white",  fillColor=blue, stroke = TRUE) %>% 
  
  ### add nh asian stops
  addCircleMarkers(data = nh_asian_dots, group = "Asian", label = NULL, 
                   weight = 2, opacity = 0.2, fillOpacity = 1, radius = 2, 
                   color="white",  fillColor=  halloween_orange, stroke = TRUE) %>% 
    
  ### add sswana stops
  addCircleMarkers(data = sswana_dots, group = "SWANA/SA", label = NULL, 
                   weight = 2, opacity = 0.2, fillOpacity = 1, radius = 2, 
                   color="white",  fillColor= american_yellow, stroke = TRUE) %>% 
    
  ### add nh_nhpi stops
  addCircleMarkers(data = nh_nhpi_dots, group = "NHPI", label = NULL, 
                   weight = 2, opacity = 0.2, fillOpacity = 1, radius = 2, 
                   color="white",  fillColor= halaya_ube, stroke = TRUE) %>% 
    
  ### add NH AIAN stops
  addCircleMarkers(data = nh_aian_dots,  group = "AIAN", label = NULL, 
                   weight = 2, opacity = 0.2, fillOpacity = 1, radius = 2, 
                   color="white",  fillColor= dark_tangerine, stroke = TRUE) %>%
    
  ### add Division polygons
  addPolygons(data = div_wide_sf, label=~htmlEscape(paste0(div_name, " Division ")),
              color = "black", fillColor = "white", opacity = 1, 
              popup = ~paste0(rate_popup), weight = 2, smoothFactor = 0.5,  
              highlight = highlightOptions(color = "white", weight = 3, 
                                           bringToFront = TRUE, 
                                           sendToBack = FALSE)) %>%
  
  ### Layer control checkbox for race/ethnic groups
  addLayersControl(overlayGroups = map_race_labels, 
                   options = layersControlOptions(collapsed = FALSE)) %>%
    
  ### add Legend
  addLegend("bottomleft", title= paste0("1 dot = 5 field interview card stops"), 
            colors = colors_function(colors = map_colors, sizes = "10", 
                                     shapes = "circle", borders = NULL),
            labels= map_race_labels, opacity = 0.7, labFormat = labelFormat(digits = 0)) %>% 
  
  ## add custom popup with full content
  addControl(html=rate_popup, position = "topleft") %>%
  
  ### set tiles and view
  addProviderTiles("CartoDB.Positron") %>%
  setView(-117.13395, 32.79923, zoom = 10) %>% #base and view 
    
  ### Deselect White from Layer Control checkbox
  hideGroup(c("White"))   # hide white group
  
```
  
<h3 class="leaflet-map-title">SDPD disproportionately conducts field interviews for people of color in the Southeastern and Central Division</h3>
<h4 class="leaflet-map-subtitle">People officers field interviewed by race and division</h4>
    
```{r, echo=FALSE}
  map
```
  <div align="left" style="line-height: 1em;"><span style=' font-size: 10px;'>Race/ethnicity: AIAN=American Indian or Alaska Native, NHPI=Native Hawaiian or Pacific Islander, SWANA/SA=Southwest Asian (Middle Eastern) or North African, or South Asian.</span> <span style="font-size: 10px;">Stops are randomly placed within each division and do not indicate the exact location of the stop.</span> <span style=' font-size: 10px;'>Catalyst California's calculations based on City of San Diego's Police Stop Data (2022); <u>[Catalyst California](https://www.catalystcalifornia.org)</u>, 2023. Analysis for all officer-initiated stops that resulted in a field interview card completed.</div>
    
    
  