---
title: "San Diego RIPA Data Pillars Project"
subtitle: "Descriptive Analysis"
author: "Catalyst California"
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
    toc: yes
    toc_depth: 2
    toc_float: yes
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)


list.of.packages <- c("usethis", "highcharter", "tidyverse", "sf", "leaflet", "sp", "htmltools", "magrittr", "flextable", "htmlwidgets")

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

#### Load libraries ####
library(usethis)
library(highcharter)
library(tidyverse)
library(sf)
library(leaflet)
library(sp)
library(htmltools) # for html escape
library(magrittr)
library(flextable)
library(htmlwidgets)
library(sf)
library(leaflet)
library(sp)
library(htmltools) # for html escape
library(scales)

#add function that better fits the tables to the page    
    FitFlextableToPage <- function(ft, pgwidth = 6.5){

  ft_out <- ft %>% autofit()

  ft_out <- width(ft_out, width = dim(ft_out)$widths*pgwidth /(flextable_dim(ft_out)$widths))
  return(ft_out)
    }

#### Add database connection ####

source("W:\\RDA Team\\R\\credentials_source.R")
con <- connect_to_db("rjs_pillars")

# read in tables from postgres

stops<-dbGetQuery(con, "SELECT * FROM rel_stops") 

# Add column just for year of stop to stops table
stops$Year = substr(stops$date_stop, 1, 4)


reason<-dbGetQuery(con, "SELECT * FROM rel_stopreason")
person<-dbGetQuery(con, "SELECT * FROM rel_persons")
result<-dbGetQuery(con, "SELECT * FROM rel_stopresult")
search<-dbGetQuery(con, "SELECT * FROM rel_basis_search")
contra<-dbGetQuery(con, "SELECT * FROM rel_contrabands_evidences")
race<-dbGetQuery(con, "SELECT * FROM rel_races_recode")
beats<-st_read(con, query = "SELECT * FROM sangis_sdpd_beats_2023")%>%
  mutate(beat=as.character(beat))
div<-dbGetQuery(con, "SELECT * FROM sangis_sdpd_divisions_2023")%>%
  select(-geom)


```

Before engaging in a detailed analysis of SDPD patrol activity as outlined in the research plan, we created this descriptive analysis to explore overall trends in the SDPD RIPA data. This helps us determine if our expected findings are likely to show up in the data and allows us to identify any trends we did not expect. The analysis includes: 

General counts and percents of SDPD stops by key fields in RIPA data, e.g., stop reason, stop result, calls for service, race, age, searches, officer assignment. 

# Total Number of Stops and People Stopped

In 2022, there were __86,632 unique police stops__. Of those stops, __79, 416 were officer-initiated stops.__ Gang enforcement made __420 unique stops__ in 2022. Of those stops, __397 were officer-initiated stops__. 

In 2022, __96,119 people__ were stopped by SDPD, __88,099__ of which were stopped for an officer-initiated stop. Gang enforcement stopped __508 people__ in 2022; __484 people__ were stopped for an officer-initiated stop.  


```{r}

stop_tot<-stops%>%
 filter(Year=='2022')%>%
  group_by(stop_id)%>%
  summarise(count=n())%>%
  mutate(total=sum(count))

stop_tot_cfs<-stops%>%
 filter(Year=='2022' & stop_in_response_to_cfs == 0)%>%
  group_by(stop_id)%>%
  summarise(count=n())%>%
  mutate(total=sum(count))

stop_tot_gu<-stops%>%
 filter(Year=='2022' & assignment =='Gang enforcement')%>%
  group_by(stop_id)%>%
  summarise(count=n())%>%
  mutate(total=sum(count))


stop_tot_gu_cfs<-stops%>%
 filter(Year=='2022' & assignment =='Gang enforcement' & stop_in_response_to_cfs == 0)%>%
  group_by(stop_id)%>%
  summarise(count=n())%>%
  mutate(total=sum(count))

ppl_tot<-person%>%
  # join stops table so we can filter for year / cfs
  left_join(stops%>%select(stop_id, Year, stop_in_response_to_cfs, assignment ))%>%
  filter(Year=='2022')%>%
  group_by(stop_id, person_id)%>%
  summarise(count=n())%>%
  ungroup()%>%
  mutate(total=sum(count))

ppl_tot_offit<-person%>%
  # join stops table so we can filter for year / cfs
  left_join(stops%>%select(stop_id, Year, stop_in_response_to_cfs, assignment ))%>%
  filter(Year=='2022' & stop_in_response_to_cfs==0)%>%
  group_by(stop_id, person_id)%>%
  summarise(count=n())%>%
  ungroup()%>%
  mutate(total=sum(count))


ppl_tot_gu<-person%>%
  # join stops table so we can filter for year / cfs
  left_join(stops%>%select(stop_id, Year, stop_in_response_to_cfs, assignment ))%>%
  filter(Year=='2022' & assignment == 'Gang enforcement')%>%
  group_by(stop_id, person_id)%>%
  summarise(count=n())%>%
  ungroup()%>%
  mutate(total=sum(count))


ppl_tot_gu_offit<-person%>%
  # join stops table so we can filter for year / cfs
  left_join(stops%>%select(stop_id, Year, stop_in_response_to_cfs, assignment ))%>%
  filter(Year=='2022' & assignment == 'Gang enforcement' & stop_in_response_to_cfs==0)%>%
  group_by(stop_id, person_id)%>%
  summarise(count=n())%>%
  ungroup()%>%
  mutate(total=sum(count))

```


# Calls for Service versus Officer-initiated

The majority of officer time is spent on officer-initiated stops rather than calls for service. This is true across all officer unit types, as well as for the gang enforcement unit.

Analysis cut by all unit types:

```{r}

# prep data 

cfs<-stops%>%
  select(stop_id, Year, stop_in_response_to_cfs, stopduration)%>%
  filter(Year=='2022')%>%
  mutate(Total=n())%>%
  group_by(stop_in_response_to_cfs)%>%
  mutate(Count=n(),
         Percent=Count/Total*100,
         'Average Stop Time (Mins)' = mean(stopduration, na.rm = FALSE),
            'Median Stop Time (Mins)'=   median(stopduration,na.rm = FALSE),
             'Min Stop Time (Mins)'=min(stopduration,na.rm = FALSE),
            'Max Stop Time (Mins)'=max(stopduration,na.rm = FALSE),
             'Total Time (Mins)'= sum(stopduration)
         )%>%
  slice(1)%>%
  select(-c(stop_id, stopduration))%>%
  mutate(stop_in_response_to_cfs=ifelse(stop_in_response_to_cfs %in% 0, 'No', 'Yes'))%>%
  rename('Call for Service'='stop_in_response_to_cfs')%>%
  mutate(Assignment = 'All units')%>%
  relocate(Assignment, .after = Year)
         


#create and style flextable

  flextable(cfs) %>% 
colformat_double(j = c("Percent", "Average Stop Time (Mins)",'Median Stop Time (Mins)',
                       'Min Stop Time (Mins)','Max Stop Time (Mins)', 'Total Time (Mins)'), digits = 1)%>% 
  set_caption('Stop Counts and Rates by Response to Service Call for All Units') %>% 
  theme_vanilla() 

```

Analysis filtered for gang suppression unit:

```{r}

# prep data 

cfs_gu<-stops%>%
  select(stop_id, Year, assignment, stop_in_response_to_cfs, stopduration)%>%
  filter(Year=='2022' & assignment =='Gang enforcement')%>%
  mutate(Total=n())%>%
  group_by(stop_in_response_to_cfs)%>%
  mutate(Count=n(),
         Percent=Count/Total*100,
         'Average Stop Time (Mins)' = mean(stopduration, na.rm = FALSE),
            'Median Stop Time (Mins)'=   median(stopduration,na.rm = FALSE),
             'Min Stop Time (Mins)'=min(stopduration,na.rm = FALSE),
            'Max Stop Time (Mins)'=max(stopduration,na.rm = FALSE),
             'Total Time (Mins)'= sum(stopduration)
         )%>%
  slice(1)%>%
  select(-c(stop_id, stopduration))%>%
  mutate(stop_in_response_to_cfs=ifelse(stop_in_response_to_cfs %in% 0, 'No', 'Yes'))%>%
  rename('Call for Service'='stop_in_response_to_cfs',
         'Assignment'='assignment')
         


#create and style flextable

  flextable(cfs_gu) %>% 
colformat_double(j = c("Percent", "Average Stop Time (Mins)",'Median Stop Time (Mins)',
                       'Min Stop Time (Mins)','Max Stop Time (Mins)', 'Total Time (Mins)'), digits = 1)%>% 
  set_caption('Stop Counts and Rates by Response to Service Call for Gang Suppression Unit') %>% 
  theme_vanilla() 

```

Analysis cut by SDPD division:

All SDPD divisions spent more than 90% of its time on officer-initiated stops, with the exception of Mid-City and Eastern which spent over 87% of patrol time on officer-initiated stops.

```{r}

# join divisions to the beats table just to get the full division name

beats_div<-beats%>%
  left_join(div%>%select(div_num, div_name), by=c("div"="div_num"))%>%
  select(beat, name, div, div_name)%>%
    as.data.frame()%>%
  select(-geom)%>%
  rename('beat_name' = 'name')%>%
  group_by(beat)%>%
  slice(1)%>%
  mutate(beat_name=ifelse(beat %in% "511", 'BARRIO LOGAN', beat_name))%>% # there was a data entry error where beat=511 showed up twice, once with the beat_name included, and once where beat_name was NA. So I manually re-added the beat_name
  ungroup()

# join updated beats-div crosswalk to stops data

beats_stop<-stops%>%
  left_join(beats_div%>%select(beat, div, div_name), by=c("beat"="beat"))

# analyze data 

# explore stops where div = NA. These look to be where beat = 999, Unknown 999 but lets confirm

na<-beats_stop%>%
  select(stop_id, Year, div_name, stop_in_response_to_cfs, stopduration)%>%
  filter(Year=='2022')%>%
  group_by(div_name)%>%
  mutate(Total=n())%>%
  filter(is.na(div_name))%>%
  left_join(stops%>%select(stop_id, beat))

# table(na$beat) # there are mostly beat=999, which is unknown. There is also no beat=125 in the sdpd beats df. I think it is ok to recode these all to just 'unknown beat' 

# move on to main analysis 

df<-beats_stop%>%
  select(stop_id, Year, div_name, stop_in_response_to_cfs, stopduration)%>%
  filter(Year=='2022')%>%
  group_by(div_name)%>%
  mutate(Total=n())%>%
  mutate(div_name=ifelse(is.na(div_name), 'UNKNOWN DIVISION', div_name))%>%
  group_by(div_name,stop_in_response_to_cfs)%>%
  mutate(Count=n(),
         Percent=Count/Total*100,
         'Average Stop Time (Mins)' = mean(stopduration, na.rm = FALSE),
            'Median Stop Time (Mins)'=   median(stopduration,na.rm = FALSE),
             'Min Stop Time (Mins)'=min(stopduration,na.rm = FALSE),
            'Max Stop Time (Mins)'=max(stopduration,na.rm = FALSE),
             'Total Time (Mins)'= sum(stopduration)
         )%>%
  slice(1)%>%
  select(-c(stop_id, stopduration))%>%
  mutate(stop_in_response_to_cfs=ifelse(stop_in_response_to_cfs %in% 0, 'No', 'Yes'))%>%
  rename('Call for Service'='stop_in_response_to_cfs')%>%
  rename('Division Name' = 'div_name')



#create and style flextable

  flextable(df) %>% 
colformat_double(j = c("Percent", "Average Stop Time (Mins)",'Median Stop Time (Mins)',
                       'Min Stop Time (Mins)','Max Stop Time (Mins)', 'Total Time (Mins)'), digits = 1)%>% 
  set_caption('Stop Counts and Rates by Response to Service Call by Division') %>% 
  theme_vanilla() 



```

## Total Officer-Intiated Stops Over Time

The number of officer-initiated stops across all police unit types peaked in 2019, with a slight decrease in 2020-2022. The COVID-19 pandemic directly impacted police activity, making it unsurprising that there were less stops conducted overall.

The gang suppression unit has conducted less officer-initiated stops overtime, with a steep drop-off between 2021-2022. Notably, the gang enforcement unit made more officer-initiated stops in 2020 than in 2021 and 2022 combined. This suggests that the numbers are not driven by the COVID-19 pandemic but other reasons. Officers making gang enforcement stops might be doing it under the guise of a different unit assignment, resulting in the gang enforcement stops being under-counted. 

```{r}

# calculate total number of officer initiated stops across all units

count<-stops%>%
  filter(stop_in_response_to_cfs==0)%>%
  select(Year, stop_id)%>%
  group_by(Year)%>%
  mutate('All Units Total'=n())%>%
  slice(1)%>%
  select(-stop_id)

# calculate total number of officer initiated stops for gang suppresion unit

count_gu<-stops%>%
  filter(stop_in_response_to_cfs==0 )%>%
  filter(grepl('Gang enforcement', assignment))%>%
  select(Year, stop_id)%>%
  group_by(Year)%>%
  mutate('Gang Suppression Unit Total'=n())%>%
  slice(1)

# combine both unit tables

count<-count%>%
  left_join(count_gu%>%select(Year, 'Gang Suppression Unit Total'))
  
  #create and style flextable

  flextable(count) %>% 
  set_caption('Total Count of Officer-Initiated Stops Over Time') %>% 
  theme_vanilla() %>%
  width(width=2,unit="in")
  
  
```

## Total Officer-Intiated Stops by Police Assignment

Patrol, traffic enforcement, and field operations conduct the majority of officer-initiated stops, followed by the 'Other' unit category. While gang enforcement has the third highest rate of officer-initiated stops the overall count is low. This is most likely due to gang enforcement stops being conducted under a different unit assignment.

```{r}

units22<-stops%>%
  filter(stop_in_response_to_cfs==0)%>%
  select(Year, stop_id, officer_assignment_key, assignment)%>%
  group_by(Year)%>%
  mutate(Total=n())%>%
  group_by(Year, assignment)%>%
  summarise(Total=Total,
            Count=n(),
            Percent=Count/Total*100)%>%
  slice(1)%>%
  arrange(-Percent)%>%
  rename('Patrol Assignment'='assignment')%>%
  filter(grepl('2022', Year))


#create and style flextable

  flextable(units22) %>% 
colformat_double(j = c("Percent"), digits = 1)%>% 
  set_caption('Total and Rate of Stops by Patrol Assignment Type in 2022') %>% 
  theme_vanilla() %>%
  width(j=2,2.5,unit="in")


```

## Total Officer-Intiated Stops by Police Division

High rates of officer-initiated stops are consistent across all police divisions. 

```{r}

df<-beats_stop%>%
  filter(Year == "2022")%>%
  select(Year, stop_id, div_name, stop_in_response_to_cfs)%>%
  group_by(div_name)%>%
  mutate(Total=n())%>%
  group_by(div_name, stop_in_response_to_cfs)%>%
  summarise(Year,
            Total=Total,
            Count=n(),
            Percent=Count/Total*100)%>%
  slice(1)%>%
    mutate(div_name=ifelse(is.na(div_name), 'UNKNOWN DIVISION', div_name))%>%
  mutate(stop_in_response_to_cfs=ifelse(stop_in_response_to_cfs %in% 0, 'No', 'Yes'))%>%
  rename('Call for Service'='stop_in_response_to_cfs')%>%
  rename('Division Name'='div_name')%>%
  relocate(Year, .before = "Division Name")


#create and style flextable

  flextable(df) %>% 
colformat_double(j = c("Percent"), digits = 1)%>% 
  set_caption('Total and Rate of Stops by Police Division in 2022') %>% 
  theme_vanilla() %>%
  width(j=2,2.3,unit="in")








```


# Stop Reasons

Traffic Violation is the most common stop reason for officer-initiated stops, followed by Reasonable Suspicion, across all police units and for the gang enforcement unit. However, nearly 70% of all officer-initiated stops conducted by the gang enforcement unit are for Traffic Violation, with only 24% being conducted for Reasonable Suspicion. Traffic Violation stops are a more significant majority of all gang enforcement stops. 

## Stop reason by police assignment

```{r}

# for the rel_stopreason table, it seems like the reason why some people show up in multiple rows is because
# there are different values for 'reason_for_stop_text' or 'reason_for_stop_detail' or 'reason_stop_explaination' BUT the 'reason_for_stop' value remains the same. Which is ultimately what we want for our analysis. So my goal is to create a flattened version of rel_stopreason that shows each person once for their stop reason. UNLESS a person shows up more than once because there is more than one unique stop reason

# similarliy seeing now that rel_persons also has multiple rows for people who have more than 1 perceived age or gender


# Continuing analysis for now:

reason_p<-person%>%
  
  # join stops table so we can filter for year / cfs
  
  left_join(stops%>%select(stop_id, Year, stop_in_response_to_cfs ))%>%
  filter(Year=='2022' & stop_in_response_to_cfs == 0)%>%
  
  # left join reasons -filtered for year/cfs - to get stop reason
  
  left_join(reason%>%
              left_join(stops%>%select(stop_id, Year, stop_in_response_to_cfs ))%>%
              filter(Year=='2022' & stop_in_response_to_cfs == 0)%>%
              select(stop_id, person_id, reason_for_stop))%>%
  
  select(Year, stop_id, person_id, reason_for_stop)

# Run a quick test to see if one person/stop_id combo has more than 1 unique reason_for_stop

test<-reason_p %>%
  group_by(stop_id, person_id) %>% 
  mutate(same = +(n_distinct(reason_for_stop) == 1)) %>% 
  ungroup()

 # table(test$same) # all 90777 rows return same==1 which means there are not any people who were stopped for different reason_for_stop

# Calculate estimates -all units

reason_p<-reason_p%>%
  mutate(Total=n())%>%
  group_by(reason_for_stop)%>%
    mutate(Count=n(),
           Percent=Count/Total*100)%>%
  slice(1)%>%
   select(-c(stop_id, person_id))%>%
  rename( 'Reason for stop'='reason_for_stop')%>%
  arrange(-Percent)%>%
  mutate(Assignment='All units')%>% 
  relocate(Assignment, .after = Year)

# Calculate estimates -gang suppression unit only

reason_p_gu<-person%>%
  left_join(stops%>%select(stop_id, Year, stop_in_response_to_cfs, assignment ))%>%
  filter(Year=='2022' & stop_in_response_to_cfs == 0 & assignment == 'Gang enforcement')%>%

  left_join(reason%>%
              left_join(stops%>%select(stop_id, Year, stop_in_response_to_cfs,assignment ))%>%
              filter(Year=='2022' & stop_in_response_to_cfs == 0 & assignment == 'Gang enforcement')%>%
              select(stop_id, person_id, reason_for_stop))%>%
  
  select(Year, assignment, stop_id, person_id, reason_for_stop)%>%
  mutate(Total=n())%>%
  group_by(reason_for_stop)%>%
    mutate(Count=n(),
           Percent=Count/Total*100)%>%
  slice(1)%>%
   select(-c(stop_id, person_id))%>%
  rename( 'Reason for stop'='reason_for_stop',
          'Assignment' = 'assignment')%>%
  arrange(-Percent)

# final combine

reason_p<-reason_p%>%
  rbind(reason_p_gu)


# create table

  flextable(reason_p) %>% 
colformat_double(j = c("Percent"), digits = 1)%>% 
  set_caption('Total and Rate of Officer-Initiated Stops by Stop Reason') %>% 
  theme_vanilla() %>%
  width(j=3,3.65,unit="in")


```

## Stop reason by SDPD divisions

The majority of SDPD divisions conduct officer-initiated stops for Traffic Violation, with the second most common officer-initiated stop being for Reasonable Suspicions. There is no third stop reason that has a rate comparable to Traffic Violation or Reasonable Suspicion.

Notably, the Southeastern division disproportionately stops for Traffic Violation relative to Reasonable Suspicion, along with the Northwestern and Northeastern divisions. 

```{r fig.height = 7, fig.width = 13}

reason_div<-person%>%
  
  # join person table  to beats/stop data so we can filter for year / cfs
  
  left_join(beats_stop%>%select(stop_id, Year, stop_in_response_to_cfs, div_name))%>%
  filter(Year=='2022' & stop_in_response_to_cfs == 0)

# double check the na division names

na<-reason_div%>%
  filter(is.na(div_name))%>%
  left_join(stops%>%select(stop_id,  beat))

# table(na$beat) # check what the beat numbers are that have NA division names

# recode the NA division names and join to reasons table

reason_div<-reason_div%>%
  mutate(div_name=ifelse(is.na(div_name), 'UNKNOWN DIVISION', div_name)) %>%
  
  left_join(reason%>%
              select(stop_id, person_id, reason_for_stop))

# we can see that some people in reason_div show up more than once because, while the overarching stop_reason is the same there are unique details added. We said we are ok with keeping for now since it isn't that many and will be a slight overcount.

df<-reason_div%>%
  select(Year, div_name, reason_for_stop)%>%
  group_by(div_name)%>%
  mutate(Total=n())%>%
  group_by(div_name, reason_for_stop)%>%
  mutate(Count=n(),
         Percent=Count/Total*100)%>%
  arrange(-Percent,  .by_group = TRUE)%>%
  slice(1)%>%
  filter(grepl('Traffic Violation|Reasonable Suspicion|Consensual Encounter resulting in a search', reason_for_stop))

# graph

ggplot(df, aes( x = reason_for_stop , y =  Percent) ) + 
      geom_bar( stat = "identity", fill="#362178") + 
  geom_text(aes( label = paste0(round(Percent, digits=1),"%" ),
                 y= Percent),vjust = 1.4, size = 4, color = "white",
            fontface = "bold")+
      facet_wrap( ~ div_name, nrow=2) + 
      xlab("")+
  ylab("Percent of Officer-Initiated Stops (%)")+
    scale_x_discrete(labels = wrap_format(10))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=12),
        strip.background = element_rect(fill = "#362178"),
        strip.text = element_text(color = "white", size=12, face="bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
panel.background = element_blank())


```

# Stop Results

## Stop result by police assignment

Across all police units, Citation for Infraction and Warning (verbal or written) are the most common stop results, at 27.5% and 25.8% respectively. These two stop results make up approximately half of all stop results. 

The gang enforcement unit stop results do not mirror the trends across all police units. The most common stop result for a gang enforcement stop is Field Interview Card Completed, comprising of 40.2% of all stop results. Warning (verbal or written) has the second highest rate at 20.6%. 

```{r}

# calculate for all units

result_p<-person%>%
  
  # join result table to person table so we can filter for year / cfs
  
  left_join(stops%>%select(stop_id, Year, stop_in_response_to_cfs ))%>%
  filter(Year=='2022' & stop_in_response_to_cfs == 0)%>%
  
  # left join results -filtered for year/cfs - to get stop reason
  
  left_join(result%>%
              left_join(stops%>%select(stop_id, Year, stop_in_response_to_cfs ))%>%
              filter(Year=='2022' & stop_in_response_to_cfs == 0)%>%
              select(stop_id, person_id, result))%>%
  
  select(Year, stop_id, person_id, result)%>%
  mutate(Total=n())%>%
  group_by(result)%>%
    mutate(Count=n(),
           Percent=Count/Total*100)%>%
  slice(1)%>%
   select(-c(stop_id, person_id))%>%
  rename( 'Result of Stop'='result')%>%
  arrange(-Percent)%>%
  mutate(Assignment='All units')%>% 
  relocate(Assignment, .after = Year)

# repeat above steps for gang suppression unit

result_p_gu<-person%>%
  left_join(stops%>%select(stop_id, Year, stop_in_response_to_cfs, assignment ))%>%
  filter(Year=='2022' & stop_in_response_to_cfs == 0 & assignment == 'Gang enforcement')%>%

  left_join(result%>%
              left_join(stops%>%select(stop_id, Year, stop_in_response_to_cfs,assignment ))%>%
              filter(Year=='2022' & stop_in_response_to_cfs == 0 & assignment == 'Gang enforcement')%>%
              select(stop_id, person_id, result))%>%
  
  select(Year, assignment, stop_id, person_id, result)%>%
  mutate(Total=n())%>%
  group_by(result)%>%
    mutate(Count=n(),
           Percent=Count/Total*100)%>%
  slice(1)%>%
   select(-c(stop_id, person_id))%>%
  rename( 'Result of Stop'='result',
          'Assignment' = 'assignment')%>%
  arrange(-Percent)

# final combine

result_p<-result_p%>%
  rbind(result_p_gu)

# create table

  flextable(result_p) %>% 
colformat_double(j = c("Percent"), digits = 1)%>% 
  set_caption('Total and Rate of Officer-Initiated Stops by Stop Result') %>% 
  theme_vanilla() %>%
  width(j=3,3.65,unit="in")





```

## Stop result by SDPD divisions

Stop result trends vary across SDPD divisions. 

```{r fig.height = 7, fig.width = 13}

result_div<-person%>%
  
  # join person table to beats/stops data so we can filter for year / cfs
  
  left_join(beats_stop%>%select(stop_id, Year, stop_in_response_to_cfs, div_name))%>%
  filter(Year=='2022' & stop_in_response_to_cfs == 0)

# double check the na division names

na<-result_div%>%
  filter(is.na(div_name))%>%
  left_join(stops%>%select(stop_id,  beat))

# table(na$beat) # check what the beat numbers are that have NA division names

# recode the NA division names and join to reasons table. We know that some people show up more than once (multiple rows) when joining the person-results table together because they are stopped and there was more than one result. Example: stop_id=504045. We are OK for now with the slight overcount. You can see that instead of 88,099 rows after joining the results table you get 95,543 because of the multiple results per person for some. 

result_div<-result_div%>%
  mutate(div_name=ifelse(is.na(div_name), 'UNKNOWN DIVISION', div_name)) %>%
  
  left_join(result%>%
              select(stop_id, person_id, result))

# analyze

df<-result_div%>%
  select(Year, div_name, result)%>%
  group_by(div_name)%>%
  mutate(Total=n())%>%
  group_by(div_name, result)%>%
  mutate(Count=n(),
         Percent=Count/Total*100)%>%
  arrange(-Percent,  .by_group = TRUE)%>%
  slice(1)%>%
  mutate(result=ifelse(result %in% 'Warning (verbal or written)','Warning',result))%>% # lets make the warning label shorter
  filter(grepl('No Action|Field interview card completed|Citation for infraction|Warning', result))

# graph

ggplot(df, aes( x = result , y =  Percent) ) + 
      geom_bar( stat = "identity", fill="#362178") + 
  geom_text(aes( label = paste0(round(Percent, digits=1),"%" ),
                 y= Percent),vjust = 1.4, size = 4, color = "white",
             fontface = "bold")+
      facet_wrap( ~ div_name, nrow=2) + 
      xlab("")+
  ylab("Percent of Officer-Initiated Stops (%)")+
    scale_x_discrete(labels = wrap_format(20))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=15),
        strip.background = element_rect(fill = "#362178"),
        strip.text = element_text(color = "white", size=12, face="bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
panel.background = element_blank())



```

## Hit Rates

This analysis looks at what percentage of stops that resulted in a search produced contraband. Approximately 70% of all stops that result in a search do not produce any contraband across all police units and for gang enforcement stops. 

```{r}

#### All units analysis ####

# Take searches, filter for year and cf==0 and limit only to where searches took place

search_p<- search%>%
  left_join(stops%>%select(stop_id, Year, stop_in_response_to_cfs ))%>%
  filter(Year=='2022' & stop_in_response_to_cfs == 0)%>%
mutate_all(na_if,"")%>%
   filter(!is.na(basis_for_search))%>%
  select(Year, stop_id, person_id, basis_for_search)%>%
    group_by(stop_id, person_id)%>%
  slice(1)%>%
  ungroup()%>%
  mutate(search_total=n())

# Grab  contraband and filter for year and where cfs ==0, and perform slice=1 to get only 1 person per row. So if person had multiple contraband found, have them only show up once

contra_p<-contra%>%
  left_join(stops%>%select(stop_id, Year, stop_in_response_to_cfs ))%>%
  filter(Year=='2022' & stop_in_response_to_cfs == 0)%>%
  select(Year, stop_id, person_id, contraband)%>%
    group_by(stop_id, person_id)%>%
slice(1)

# now focus on those who had NO contraband found, need to join with search data because we need to exclude people who had no contraband found AND no search. We want to be left only with people who were searched + no contraband found

contra_none<-contra_p%>%
  filter(contraband=='None')%>%
  left_join(search_p)%>%
  filter(!is.na(basis_for_search))%>%
  ungroup()%>%
  mutate(Percent=n()/search_total*100)%>%
  mutate(Year='2022',
         Assignment='All units')%>%
  rename('Contraband'='contraband',
         'Search Total'='search_total') %>%
  filter(row_number()==1)%>%
  select(Year, Assignment, Contraband, 'Search Total', Percent)

# Now lets focus on finding the rate of searches where 1 or more contraband were found

# again, first join contraband to searches to remove instances where people were NOT searched but contraband was found

contra_p<-contra_p%>%
  filter(contraband!='None')%>%
  left_join(search_p)%>%
  filter(!is.na(basis_for_search))%>%
  ungroup()%>%
  mutate(Percent=n()/search_total*100)%>%
  mutate(Year='2022',
         Assignment='All units',
         contraband=ifelse(contraband %in% 'Cell phone(s) or electronic device(s)', 'One or more contraband items', contraband))%>%
  rename('Contraband'='contraband',
         'Search Total'='search_total') %>%
  filter(row_number()==1)%>%
  select(Year, Assignment, Contraband, 'Search Total', Percent)
  
# join both tables together

hitrate_allunit<-rbind(contra_none, contra_p)

  
#### Gang suppression unit ####

# Take searches, filter for assignment,  year and cf==0 and limit only to where searches took place

search_gu<- search%>%
  left_join(stops%>%select(stop_id, Year, assignment, stop_in_response_to_cfs ))%>%
  filter(Year=='2022' & stop_in_response_to_cfs == 0 & assignment == 'Gang enforcement')%>%
mutate_all(na_if,"")%>%
   filter(!is.na(basis_for_search))%>%
  select(Year, stop_id, person_id, basis_for_search)%>%
    group_by(stop_id, person_id)%>%
  slice(1)%>%
  ungroup()%>%
  mutate(search_total=n())

# Grab  contraband and filter for year and where cfs ==0, and perform slice=1 to get only 1 person per row. So if person had multiple contraband found, have them only show up once

contra_gu<-contra%>%
  left_join(stops%>%select(stop_id, Year, assignment, stop_in_response_to_cfs ))%>%
  filter(Year=='2022' & stop_in_response_to_cfs == 0 & assignment == 'Gang enforcement')%>%
  select(Year, assignment, stop_id, person_id, contraband)%>%
    group_by(stop_id, person_id)%>%
slice(1)

# now focus on those who had NO contraband found, need to join with search data because we need to exclude people who had no contraband found AND no search. We want to be left only with people who were searched + no contraband found

contra_none<-contra_gu%>%
  filter(contraband=='None')%>%
  left_join(search_gu)%>%
  filter(!is.na(basis_for_search))%>%
  ungroup()%>%
  mutate(Percent=n()/search_total*100)%>%
  rename('Contraband'='contraband',
         'Search Total'='search_total',
         'Assignment'='assignment') %>%
  filter(row_number()==1)%>%
  select(Year, Assignment, Contraband, 'Search Total', Percent)

# Now lets focus on finding the rate of searches where 1 or more contraband were found

# again, first join contraband to searches to remove instances where people were NOT searched but contraband was found

contra_gu<-contra_gu%>%
  filter(contraband!='None')%>%
  left_join(search_gu)%>%
  filter(!is.na(basis_for_search))%>%
  ungroup()%>%
  mutate(Percent=n()/search_total*100)%>%
  mutate(contraband=ifelse(contraband %in% 'Firearm(s)',  'One or more contraband items', contraband))%>%
  rename('Contraband'='contraband',
         'Search Total'='search_total',
         "Assignment"="assignment") %>%
  filter(row_number()==1)%>%
  select(Year, Assignment, Contraband, 'Search Total', Percent)
  
# join both tables together

hitrate_gu<-rbind( contra_none, contra_gu)

#### Final combine and visualize ####

df<-rbind(hitrate_allunit, hitrate_gu)

  flextable(df) %>% 
colformat_double(j = c("Percent"), digits = 1)%>% 
  set_caption('Hit Rate of Stops that Result in Searches') %>% 
  theme_vanilla() %>%
  width(j=3,2,unit="in")


```

# Age Demographics

## Stops by age by poliice assignment 

The gang enforcement unit stops more individuals between the age of 18-34 compared to all police units. 

```{r}
# Note that the numbers might be slightly inflated because there are some people that are indicated as having more than 1 perceived age. But for the purposes of this descriptive analysis we want to see general trends and are OK with some duplicates for the time being. 


# take persons filter for year/cfs, and create age brackets out of perceived age

person_age<-person%>%
  left_join(stops%>%select(stop_id, Year, assignment, stop_in_response_to_cfs ))%>%
  filter(Year=='2022' & stop_in_response_to_cfs == 0)%>%
  select(stop_id, person_id, Year, assignment, perceived_age)%>%
  mutate(age_bracket=ifelse(perceived_age <=18, '18 and under',
                      ifelse(perceived_age <=24 & perceived_age>18, '19-24',
                       ifelse(perceived_age <=34 & perceived_age>24, '25-34',
                        ifelse(perceived_age <=44 & perceived_age>34, '35-44',
                       ifelse(perceived_age <=54 & perceived_age>44, '45-54',
                        ifelse(perceived_age <=64 & perceived_age>54, '55-64',
                                ifelse(perceived_age >=65, '65 and older', 'Blank'))))))))

# calculate rates for all units

  person__age_allunit<-person_age%>%
    mutate(Total=n())%>%
    group_by(age_bracket)%>%
      mutate(Count=n(),
           Percent=Count/Total*100)%>%
  slice(1)%>%
   select(-c(stop_id, person_id, perceived_age, assignment))%>%
  rename( 'Age'='age_bracket')%>%
  # arrange(-Percent)%>%
  mutate(Assignment='All units')%>% 
  relocate(Assignment, .after = Year)
  
  # calculate rate for gang suppression unit only

   person_age_gu<-person_age%>%
     filter(assignment == 'Gang enforcement')%>%
    mutate(Total=n())%>%
    group_by(age_bracket)%>%
      mutate(Count=n(),
           Percent=Count/Total*100)%>%
  slice(1)%>%
   select(-c(stop_id, person_id, perceived_age))%>%
  rename( 'Age'='age_bracket',
          'Assignment'='assignment')
  
  
# final combine
   
   df<-rbind(person__age_allunit, person_age_gu)
   
# create table

  flextable(df) %>% 
colformat_double(j = c("Percent"), digits = 1)%>% 
  set_caption('Total and Rate of Officer-Initiated Stops by Age Group') %>% 
  theme_vanilla() %>%
  width(j=2,1.5,unit="in")



```

## Stops by age by SDPD divisions

The Southeastern and Northern divisions disproportionately stops individuals age 25-34 relative to other age groups. 

```{r fig.height = 7, fig.width = 13}

# take the person_age table made in the previous code chunk already filtered for year/cfs and recoded age labels. Join to beats_stops to get division names

age_div<-person_age%>%
  left_join(beats_stop%>%select(stop_id, Year, stop_in_response_to_cfs, div_name))
  
# calculate rates 

  df<-age_div%>%
    group_by(div_name)%>%
    mutate(Total=n())%>%
    group_by(div_name, age_bracket)%>%
      mutate(Count=n(),
           Percent=Count/Total*100)%>%
  slice(1)
  
# graph

ggplot(df, aes( x = age_bracket , y =  Percent) ) + 
      geom_bar( stat = "identity", fill="#362178") + 
  geom_text(aes( label = paste0(round(Percent, digits=1),"%" ),
                 y= Percent),vjust = 1.4, size = 3, color = "white",
             fontface = "bold")+
      facet_wrap( ~ div_name, nrow=2) + 
      xlab("")+
  ylab("Percent of Officer-Initiated Stops (%)")+
    scale_x_discrete(labels = wrap_format(10))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=15),
        strip.background = element_rect(fill = "#362178"),
        strip.text = element_text(color = "white", size=12, face="bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
panel.background = element_blank())

  
  

```

# Race/Ethnicity Demographics

## Stops by race by police assignment

The gang enforcement unit disproportionately stops Black and Latinx individuals. 

```{r}

#### All Units level analysis ####

# calculate rates for all units: nh_race 

  race_allunit_nh<-race%>%
  left_join(stops%>%select(stop_id, Year, assignment, stop_in_response_to_cfs ))%>%
  filter(Year=='2022' & stop_in_response_to_cfs == 0)%>%
  select(Year, nh_race)%>%
    mutate(Total=n())%>%
    group_by(nh_race)%>%
      mutate(Count=n(),
           Percent=Count/Total*100)%>%
  slice(1)%>%
  rename( 'Race'='nh_race')%>%
  arrange(-Percent)%>%
  mutate(Assignment='All units')%>% 
  relocate(Assignment, .after = Year)%>%
  filter(!grepl('nh_mesa|nh_aian|nh_nhpi', Race))%>%
  drop_na(Race) # drop the row where nh_race = NA because that is for the people with 5+ perceived number of races who were recoded tonh_race = NULL

# calculate rates for all units: mesa/aian/nhpi

 race_allunit_other<-race%>%
  left_join(stops%>%select(stop_id, Year, assignment, stop_in_response_to_cfs ))%>%
  filter(Year=='2022' & stop_in_response_to_cfs == 0)%>%
  select(Year, mesa_flag, aian_flag, nhpi_flag)%>%
    mutate(Total=n())%>%
  summarise(Year, 
            Total,
            mesa=sum(mesa_flag, na.rm=TRUE),
            nhpi=sum(nhpi_flag, na.rm=TRUE),
            aian=sum(aian_flag, na.rm=TRUE))%>%
   pivot_longer(3:5, names_to = "Race", values_to="Count" )%>%
   group_by(Race)%>%
   slice(1)%>%
   mutate(Percent=Count/Total*100)%>%
   mutate(Assignment='All units')
 
 
 # fix order of columns
 race_allunit_other<-race_allunit_other[,c(1,6,3,2,4,5)]
 
 # combine: 
 
 race_allunit<-rbind(race_allunit_nh, race_allunit_other)%>%
   arrange(-Percent)
 
 #### Gang Suppression Unit only analysis ####
 
 # calculate rate/count for nh_race only for gang suppression unit

  race_gu_nh<-race%>%
  left_join(stops%>%select(stop_id, Year, assignment, stop_in_response_to_cfs ))%>%
  filter(Year=='2022' & stop_in_response_to_cfs == 0 & assignment == 'Gang enforcement')%>%
  select(Year, nh_race, assignment)%>%
    mutate(Total=n())%>%
    group_by(nh_race)%>%
      mutate(Count=n(),
           Percent=Count/Total*100)%>%
  slice(1)%>%
  rename( 'Race'='nh_race')%>%
  arrange(-Percent)%>%
   rename('Assignment'='assignment')%>%
     relocate(Assignment, .after = Year)%>%
  filter(!grepl('nh_mesa|nh_aian|nh_nhpi', Race))%>%
  drop_na(Race) # drop the row where nh_race = NA because that is for the people with 5+ perceived number of races who were recoded tonh_race = NULL
 
 
 # calculate rates for all units: mesa/aian/nhpi

 race_gu_other<-race%>%
  left_join(stops%>%select(stop_id, Year, assignment, stop_in_response_to_cfs ))%>%
 filter(Year=='2022' & stop_in_response_to_cfs == 0 & assignment == 'Gang enforcement')%>%  select(Year, assignment, mesa_flag, aian_flag, nhpi_flag)%>%
    mutate(Total=n())%>%
  summarise(Year, 
            assignment,
            Total,
            mesa=sum(mesa_flag, na.rm=TRUE),
            nhpi=sum(nhpi_flag, na.rm=TRUE),
            aian=sum(aian_flag, na.rm=TRUE))%>%
   pivot_longer(4:6, names_to = "Race", values_to="Count" )%>%
   group_by(Race)%>%
   slice(1)%>%
   mutate(Percent=Count/Total*100)%>%
   rename('Assignment'='assignment')
 
  # fix order of columns
 race_gu_other<-race_gu_other[,c(1,2,4,3,5,6)]
 
 # combine:
 
 race_gu<-rbind(race_gu_nh, race_gu_other)%>%
   arrange(-Percent)
 
 #### Final combine and visualize table ####
 
 # combine all units + gang suppression unit tables and create race labels
 
 df<-rbind(race_allunit, race_gu)%>%
   mutate(Race=ifelse(Race %in% "nh_white", "White NH",
                 ifelse(Race %in% "nh_black", "Black NH",   
                        ifelse(Race %in% "latinx", "Latinx",
                               ifelse(Race %in% "nh_asian", "Asian NH",
                                      ifelse(Race %in% "mesa", "SSWANA",
                                             ifelse(Race %in% "aian", "AIAN",
                                                    ifelse(Race %in% "nhpi", "NHPI", "Multiracial"))))))))
 
 # create table

  flextable(df) %>% 
colformat_double(j = c("Percent"), digits = 1)%>% 
  set_caption('Total and Rate of Officer-Initiated Stops by Race') %>% 
  theme_vanilla() %>%
  width(j=2,1.5,unit="in")


```

## Stops by race by SDPD divisions

The Southeastern division disproportionately stops Black and Latinx individuals relative to other racial groups. The Souther division disproportionately stops Latinx individuals relative to other racial groups. 

```{r fig.height = 7, fig.width = 13}

# calculate rates for nh_race 

  div_race_nh<-race%>%
  left_join(beats_stop%>%select(stop_id, Year, beat, div_name, stop_in_response_to_cfs ))%>%
  filter(Year=='2022' & stop_in_response_to_cfs == 0)

# explore NA div names

na<-div_race_nh%>%
  filter(is.na(div_name))

# table(na$beat) # what we've been seeing consistently

# continue analysis and recode NA div names

div_race_nh<-div_race_nh%>%
  select(div_name, nh_race)%>%
  mutate(div_name=ifelse(is.na(div_name), "UNKNOWN DIVISION", div_name))%>%
  group_by(div_name)%>%
    mutate(Total=n())%>%
    group_by(div_name, nh_race)%>%
      mutate(Count=n(),
           Percent=Count/Total*100)%>%
  slice(1)%>%
  rename( 'Race'='nh_race')%>%
  filter(!grepl('nh_mesa|nh_aian|nh_nhpi', Race))%>%
  drop_na(Race) # drop the row where nh_race = NA because that is for the people with 5+ perceived number of races who were recoded to nh_race = NULL


# calculate rates for: mesa/aian/nhpi

 div_race_other<-race%>%
  left_join(beats_stop%>%select(stop_id, Year, div_name, stop_in_response_to_cfs ))%>%
  filter(Year=='2022' & stop_in_response_to_cfs == 0)%>%
  select(div_name, mesa_flag, aian_flag, nhpi_flag)%>%
     mutate(div_name=ifelse(is.na(div_name), "UNKNOWN DIVISION", div_name))%>%
   group_by(div_name)%>%
    mutate(Total=n())%>%
  summarise( Total,
            mesa=sum(mesa_flag, na.rm=TRUE),
            nhpi=sum(nhpi_flag, na.rm=TRUE),
            aian=sum(aian_flag, na.rm=TRUE))%>%
   pivot_longer(3:5, names_to = "Race", values_to="Count" )%>%
   group_by(div_name, Race)%>%
   slice(1)%>%
   mutate(Percent=Count/Total*100)
 

 
 # combine: 
 
 df<-rbind(div_race_nh, div_race_other)%>%
   arrange(-Percent)
 
 # modify race labels for graphing
 
 df<-df%>%
   mutate(Race=ifelse(Race %in% "nh_white", "White NH",
                 ifelse(Race %in% "nh_black", "Black NH",   
                        ifelse(Race %in% "latinx", "Latinx",
                               ifelse(Race %in% "nh_asian", "Asian NH",
                                      ifelse(Race %in% "mesa", "SSWANA",
                                             ifelse(Race %in% "aian", "AIAN",
                                                    ifelse(Race %in% "nhpi", "NHPI", "Multiracial"))))))))

 # graph
 
 ggplot(df, aes( x = Race , y =  Percent) ) + 
      geom_bar( stat = "identity", fill="#362178") + 
  geom_text(aes( label = paste0(round(Percent, digits=1),"%" ),
                 y= Percent),vjust = -0.4, size = 4, color ="black", fontface="bold")+
      facet_wrap( ~ div_name, nrow=2) + 
      xlab("")+
  ylab("Percent of Officer-Initiated Stops (%)")+
    scale_x_discrete(labels = wrap_format(10))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=15),
        strip.background = element_rect(fill = "#362178"),
        strip.text = element_text(color = "white", size=15, face="bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
panel.background = element_blank())

 


```

# Map of Officer-Initiated Stops

This map shows the count of officer-initiated stops in 2022 by police beat for all police units and for the gang enforcement unit.

```{r}

#### All Units Layer ####

# first summarise count of stops per beat for all units

beats_allunit<-stops%>%
   filter(Year=='2022' & stop_in_response_to_cfs == 0)%>%
  left_join(beats, by=c("beat"="beat"))%>%
  select(-20)%>%
  group_by(name)%>%
  summarise(Year, name, count=n(), geom)%>%
  slice(1)

# convert back to sf object and project for leaflet

beats_allunit<-st_as_sf(beats_allunit)%>%
  st_transform(4326)


#### Gang suppression unit layer ####

# then create the gang suppression unit layer


beats_gu<-stops%>%
   filter(Year=='2022' & stop_in_response_to_cfs == 0 & assignment == 'Gang enforcement')%>%
  left_join(beats, by=c("beat"="beat"))%>%
  select(-20)%>%
   group_by(name)%>%
  summarise(Year, name, count=n(),
             geom)%>%
  slice(1)

# convert back to sf object and project for leafjet

beats_gu<-st_as_sf(beats_gu)%>%
  st_transform(4326)


### Divisional overlay ####

# pull in divisional layer to create an overlay

  
div<-st_read(con, query = "SELECT * FROM sangis_sdpd_divisions_2023")

# transform for leaflet

div<- st_transform(div,4326)

#### Map ####

# set palette
cc_brand <- c("#E2D9FF", "#BDAFE9","#8E7ACA", "#362178","#211447")

#palette purples
pal <- colorQuantile(cc_brand,beats_allunit$count, n = 5)

pal_allunit<-colorBin(
  palette =cc_brand,
  domain = beats_allunit$count,
  bins=c(0,300,600,900,Inf),
   na.color = NA
)

pal_gu<-colorBin(
  palette =cc_brand,
  domain = beats_gu$count,
  bins=c(0,5,10,Inf),
   na.color = NA
)

# popups

popup_allunit<-paste0("Assignment: All Units","<br>","Beat Name: ",beats_allunit$name,"<br>","Stop count: ", beats_allunit$count)

popup_gu<-paste0("Assignment: Gang Suppression Unit","<br>","Beat Name: ",beats_gu$name,"<br>","Stop count: ", beats_gu$count)

div_popup<-paste0(div$div_name)

# legend label

labels_allunit<- c( "0-300", "300-600", "600-900", "Greater than 900")
labels_gu<- c( "0-5", "5-10", "Greater than 10")



# map


leaflet(width = "100%", height = "600px") %>%
  
  addMapPane("gu_pane", zIndex=485)%>%
  addMapPane("allunit_pane", zIndex=485)%>%
    addMapPane("div_pane", zIndex=485)%>%

  

  #  all-units layer
  
  addPolygons(data = beats_allunit,  label=~htmlEscape(paste0("Beat name: ", beats_allunit$name)), color = "#444444", weight = 1, smoothFactor = 0.5,  opacity = 1, fillOpacity =1,  fillColor = ~pal_allunit(beats_allunit$count),
               popup = ~paste0(popup_allunit), 
              highlight = highlightOptions(color = "white", weight = 3, bringToFront = FALSE), group = "All Units",
               options = pathOptions(pane = "allunit_pane")) %>%
  
  # gang suppression unit layer
  
    addPolygons(data = beats_gu,  label=~htmlEscape(paste0("Beat name: ",beats_gu$name)), color = "#444444", weight = 1, smoothFactor = 0.5,  opacity = 1, fillOpacity =1,  fillColor = ~pal_gu(beats_gu$count),
               popup = ~paste0(popup_gu),
              highlight = highlightOptions(color = "white", weight = 3, bringToFront = FALSE), group = "Gang Suppression Unit",
              options = pathOptions(pane = "gu_pane")) %>%
  
  # divisional overlay
  
   addPolygons(data=div, color="#CEEA01", weight = 1.5, smoothFactor = 0.5,
                label=~htmlEscape(paste0("Division name: ",div_popup)),
  
                           opacity = 1.0, fillOpacity = 0, group="SDPD Divisions",
               options = pathOptions(pane = "div_pane"))%>%


 ## add legends for stop count for each layer
  
  addLegend(position = "bottomleft", pal = pal_allunit, values = beats_allunit$count, opacity = .5, title = "All Units", na.label = NA,
            group="All Units",
             labFormat= function(type, cuts, p) {  
                                                  paste0(labels_allunit)
                             }) %>%

  addLegend(position = "bottomleft", pal = pal_gu, values = beats_gu$count, opacity = .5, title = "Gang Suppression Unit", na.label = NA, group="Gang Suppression Unit",
             labFormat= function(type, cuts, p) {  
                                                  paste0(labels_gu)
                             }) %>%

  # layer control
  
   addLayersControl(overlayGroups = c("All Units","Gang Suppression Unit", "SDPD Divisions"),
                   options = layersControlOptions(collapsed = FALSE),
                   position = 'topright')%>%
  
  
  #hide supervisorial districts by default
  hideGroup( "SDPD Divisions")%>%
  
  #base and view
  
  addProviderTiles("CartoDB.Positron")

```
